---
title: "Supplemental Materials"
author: "Anonymized for peer review"
#date: "12/21/2020"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: 
  html_document:
    toc: true
    toc_depth: 3
    toc_float: true
    code_folding: hide
---

```{r setup, include=FALSE}
#knitr::opts_chunk$set(echo = TRUE)
library(knitr)
knitr::opts_chunk$set(results = 'asis',   
                comment = NA,
                prompt  = FALSE,
                cache   = FALSE,
                echo = TRUE,
                warning=F,
                message=F)
library(summarytools)
st_options(plain.ascii = FALSE,
           style        = "rmarkdown",
           footnote     = NA,
           subtitle.emphasis = FALSE)
options(dplyr.summarise.inform=F)
```

```{r load packages and data, include=FALSE}
library(tidyverse)
library(here)
library(cowplot)
library(lme4)
library(car)
library(harrietr)
library(ggdendro)
library(ggimage)
library(dendroextras)
library(dendextend)
library(vegan)
library(corrplot)
library(ggcorrplot)
library(cluster)
library(factoextra)
source("Grid_helper_functions.R")
theme_set(theme_cowplot())
root_path <- here()

subj_dist_long <- read.csv(here(root_path,"analysis","paper_2020","processed_data","Grid_subject_distance_item_pairs.csv"))
ratings <- read.csv(here(root_path,"analysis","paper_2020","processed_data","ratings_data.csv"))
ratings_pairs <- read.csv(here(root_path,"analysis","paper_2020","processed_data","ratings_item_pairs.csv"))

```

# 1. Practice Phase 
To assess whether participants in each age group demonstrated understanding of the grid task during the practice phase, we investigated the degree to which participants consistently arranged images belonging to the same superordinate categories (vehicles: car and bus; animals: squirrel and bird) closer together in space. We computed the average distance between images belonging to the same superordinate category (vehicles or animals) for each participant, and then compared the average distances for item pairs sharing the same category to item pairs from different categories. 3-year-olds did not consistently arrange items belonging to the same superordinate category closer together in space, (paired t-test: t(20) = -0.046, p = .96). This suggests that children in this age group were not consistently sorting images according to superordinate categories and may have struggled with the task instructions. However, we included 3-year-olds in all analyses as we had no reason to exclude this group a priori. All other age groups consistently sorted images belonging to the same category closer together in the grid space (4-year-olds: t(34) = 5.26, p < .001; 5-year-olds: t(27) = 5.42, p < .001; 6-year-olds: t(22) = 8.55, p < .001; adults: t(39) = 29.10, p < .001).

3-year-olds during the practice phase did not appear to use superordinate categories to guide their similarity judgments. Rather, children in this age range appeared to approach the task with many different strategies: some did appear to use superordinate categories like animals and vehicles (N = 5), some separated each image far apart, viewing each stimulus as distinct (N=4), some appeared to focus on other perceptual features, like color, to guide their decisions (N = 5), and some appeared to sort at random (N = 7). However, our observations in this regard are subjective and future research will need to adjust the task to better understand and tease apart possibilities for how these youngest children understood the stimuli.

```{r practice, warning = FALSE}
library(broom)
library(gt)
# get average dist based on category_pair and age bin
avg_dist_cat_by_subj <- subj_dist_long %>%
  group_by(subject, Age, Gender, age_bin,age_group,sort,category_pair) %>%
  summarize(N=n(),avg_dist=mean(dist),ci_dist=qt(0.975, N-1)*sd(dist,na.rm=T)/sqrt(N))

#paired  t-test on average distance scores
#3 year olds
t1 <- t.test(filter(avg_dist_cat_by_subj,age_bin=="3 to 4"& sort=="Practice"&category_pair=="between")$avg_dist,
       filter(avg_dist_cat_by_subj,age_bin=="3 to 4"& sort=="Practice"&category_pair=="within")$avg_dist,
       paired=T)
#4 year olds
t2 <-t.test(filter(avg_dist_cat_by_subj,age_bin=="4 to 5"& sort=="Practice"&category_pair=="between")$avg_dist,
       filter(avg_dist_cat_by_subj,age_bin=="4 to 5"& sort=="Practice"&category_pair=="within")$avg_dist,
       paired=T)
#5 year olds
t3 <-t.test(filter(avg_dist_cat_by_subj,age_bin=="5 to 6"& sort=="Practice"&category_pair=="between")$avg_dist,
       filter(avg_dist_cat_by_subj,age_bin=="5 to 6"& sort=="Practice"&category_pair=="within")$avg_dist,
       paired=T)
#6 year olds
t4 <-t.test(filter(avg_dist_cat_by_subj,age_bin=="6 to 7"& sort=="Practice"&category_pair=="between")$avg_dist,
       filter(avg_dist_cat_by_subj,age_bin=="6 to 7"& sort=="Practice"&category_pair=="within")$avg_dist,
       paired=T)
#adults
t5 <-t.test(filter(avg_dist_cat_by_subj,age_bin=="adults"& sort=="Practice"&category_pair=="between")$avg_dist,
       filter(avg_dist_cat_by_subj,age_bin=="adults"& sort=="Practice"&category_pair=="within")$avg_dist,
       paired=T)

tab <- map_df(list(t1, t2, t3, t4, t5), tidy)
tab$age = c("3-year-olds","4-year-olds","5-year-olds","6-year-olds","adults")
tab <- tab[c("age","estimate","statistic","p.value")]

tab1<- gt(tab) %>% cols_align(align="center") %>% cols_label(age = md("**Age Groups**"), estimate=md("**Estimate**"), statistic=md("**Statistic**"), p.value=md("**P-value**")) %>%   tab_header(
    title = md("Practice Phase T-tests"))
tab1

```

# 2. Sorting by  Emotion Categories {.tabset} 

## Continuous Age Predictor
In the manuscript we examined whether images belonging to the same or different emotion categories were more likely to be placed near one another. Age was binned into 4 age groups. In the following analysis, we show sorting by emotion category across child age (as a continuous predictor) with a LOESS regression fit. We see a qualitative shift around 5 years of age in sorting by emotion category.

```{r}
avg_dist_diff_cat_by_subj <- subj_dist_long %>%
  group_by(subject, Age, Gender, age_bin,age_group,sort,category_pair) %>%
  summarize(avg_dist=mean(dist,na.rm=T))  %>%
  ungroup() %>%
  group_by(subject, Age, Gender, age_bin,age_group,sort) %>%
  pivot_wider(names_from=category_pair,values_from=avg_dist) %>%
  mutate(avg_dist_diff=between-within) %>%
  mutate(sort_name=ifelse(sort=="Sort1","Same Individual",ifelse(sort=="Sort2","Different Individuals","Practice"))) %>%
  ungroup()

ggplot(filter(avg_dist_diff_cat_by_subj,age_group=="kids"&sort!="Practice"),aes(Age,avg_dist_diff))+
  geom_point()+
  geom_smooth(size=2)+
  geom_hline(yintercept=0)+
  facet_wrap(~sort_name)+
  ylab("Sorting by Emotion Categories \nLow<------------------------>High")+
  xlab("Age (in years)")
```


## Basic Emotions
In the manuscript we examined whether images belonging to the same or different emotion categories were more likely to be placed near one another. However, since some individuals might object to including emotion categories like "neutral" or "calm", we ran an additional analysis where we only included basic emotions: anger, disgust, fear, happiness, sadness, and surprise. We found the same pattern of results as when all emotion categories are included.

```{r basic emotion prep, include=FALSE}
basic = c("ang","disg","fear","hap","sad","surp")
basic_df <- data_frame(basic)
basic_emo_dist_long <- subj_dist_long %>% filter(image_cat_1 %in% basic_df$basic) %>% filter(image_cat_2 %in% basic_df$basic)

#focus on differences in dist (same v. between categories)
avg_dist_diff_cat_by_subj <- basic_emo_dist_long %>%
  group_by(subject, Age, Gender, age_bin,age_group,sort,category_pair) %>%
  summarize(avg_dist=mean(dist,na.rm=T))  %>%
  ungroup() %>%
  group_by(subject, Age, Gender, age_bin,age_group,sort) %>%
  pivot_wider(names_from=category_pair,values_from=avg_dist) %>%
  mutate(avg_dist_diff=between-within) %>%
  mutate(sort_name=ifelse(sort=="Sort1","Same Individual",ifelse(sort=="Sort2","Different Individuals","Practice"))) %>%
  ungroup()
#average across subjects  by age bin
avg_dist_diff_cat_across_subj <- avg_dist_diff_cat_by_subj %>%
  select(-between,-within) %>%
  group_by(age_bin,age_group,sort,sort_name) %>%
  summarize(N=n(),average_dist=mean(avg_dist_diff,na.rm=T),ci_dist=qt(0.975, N-1)*sd(avg_dist_diff,na.rm=T)/sqrt(N))

# for facet wrap naming
age_names <- list(
  '3 to 4'="3 y/o",
  '4 to 5'="4 y/o",
  '5 to 6'="5 y/o",
  '6 to 7'="6 y/o",
  'adults'="adults")
age_labeller <- function(variable,value){
  return(age_names[value])
}
```

```{r basic emotion plot, warning=FALSE}

# Plot
p <- ggplot(filter(avg_dist_diff_cat_across_subj, sort!="Practice"),aes(x=sort_name,y=average_dist,fill=age_bin,color=age_bin, linetype=sort_name))+
  geom_errorbar(aes(ymin=average_dist-ci_dist,ymax=average_dist+ci_dist),width=0,size=1)+
  geom_point(size=2.5)+
  facet_wrap(~age_bin,nrow=1, labeller = age_labeller, strip.position = c("bottom"))+
  geom_hline(yintercept=0)+
  theme(legend.position=c(.1,.8), legend.title = element_blank(),
        legend.margin = margin(.5,.5,.5,.5,"cm"),
        legend.background = element_rect(fill="white",size=0.6,
                                         linetype="solid", colour ="black")) +
  scale_color_manual(values=c("#8c510a","#bf812d","#80cdc1","#35978f","#01665e"))+
  guides(color=FALSE) +
  guides(fill=FALSE) +
  ylab("Sorting by Emotion Categories \nLow<------------------------>High")+
  xlab("Age Group")+
  ylim(-.1,.3) +
  labs(linetype = "Sorting Phase") +
  theme(axis.line = element_blank(),
    axis.text.x  = element_blank(),
        axis.ticks.x= element_blank(),
        axis.title.x = element_text(size=20,face="bold"),
        axis.text.y =  element_text(size=18),
        axis.title.y= element_text(size=18,face="bold"),
        strip.text.x = element_text(size=18)) #+
p

```

## Emotion Category
In the manuscript we examined whether images belonging to the same or different emotion categories were more likely to be placed near one another. Here we show this analysis in more detail by showing the effect for each individual emotion category, rather than collapsing across all categories. Overall, we see similar effects across each emotion category.
```{r emo cat data prep, include=FALSE}
avg_dist_diff_cat_by_subj <- subj_dist_long %>%
  filter(sort!="Practice") %>% 
  group_by(subject, age_bin,age_group,sort, category_pair, shared_category) %>%
  summarize(avg_dist=mean(dist,na.rm=T))  %>%
  ungroup() %>%
  group_by(subject, age_bin,age_group,sort, shared_category, category_pair) %>%
  pivot_wider(names_from=category_pair,values_from=avg_dist) %>%
  mutate(sort_name=ifelse(sort=="Sort1","Same Individual","Different Individuals")) %>%
  rename(between_old=between)
  
temp <- avg_dist_diff_cat_by_subj %>%
  group_by(subject, sort) %>% 
  summarise(between = mean(between_old, na.rm=TRUE))

avg_dist_diff_cat_by_subj <- avg_dist_diff_cat_by_subj %>% 
  dplyr:::inner_join(temp, by = c("subject","sort")) %>% 
  filter(shared_category!="between") %>% 
  mutate(avg_dist_diff=between-within) %>%
  ungroup()

#average across subjects  by age bin
avg_dist_diff_cat_across_subj <- avg_dist_diff_cat_by_subj %>%
  select(-between,-within) %>%
  group_by(age_bin,age_group,sort,sort_name, shared_category) %>%
  summarize(N=n(),average_dist=mean(avg_dist_diff,na.rm=T),ci_dist=qt(0.975, N-1)*sd(avg_dist_diff,na.rm=T)/sqrt(N))

avg_dist_diff_cat_across_subj <- avg_dist_diff_cat_across_subj %>% 
  mutate(shared_category_labels = case_when(
    shared_category == "ang" ~"angry",
    shared_category == "calm" ~"calm",
    shared_category == "disg" ~"disgust",
    shared_category == "exc" ~"excited",
    shared_category == "fear" ~"fear",
    shared_category == "hap" ~"happy",
    shared_category == "neut" ~"neutral",
    shared_category == "sad" ~"sad",
    shared_category == "surp" ~"surprise")) %>% 
  mutate(age_bin = case_when(
    age_bin == "3 to 4" ~ "3 y/o",
    age_bin == "4 to 5" ~ "4 y/o",
    age_bin == "5 to 6" ~ "5 y/o",
    age_bin == "6 to 7" ~ "6 y/o",
    age_bin == "adults" ~ "adults"))

```


```{r emo cat plot, warning=FALSE}
p <- ggplot(avg_dist_diff_cat_across_subj,aes(x=age_bin,y=average_dist,fill=age_bin,color=age_bin, linetype=sort_name))+
  geom_errorbar(aes(ymin=average_dist-ci_dist,ymax=average_dist+ci_dist),width=0,size=1, position=position_dodge(width=0.5))+
  geom_point(size=1.5, position=position_dodge(width=0.5))+
  facet_wrap(~shared_category_labels, nrow=3, strip.position = c("top"))+
  geom_hline(yintercept=0)+
  scale_color_manual(values=c("#8c510a","#bf812d","#80cdc1","#35978f","#01665e"))+
  ylab("Sorting by Emotion Category \nLow<------------------------>High")+
  xlab("Age Group")+
  labs(linetype = "Sorting Phase:", color = "Age:", fill = "Age:") +
  theme(legend.position="top", legend.box = "vertical", legend.justification = "center") +
  theme(axis.line = element_blank(),
    axis.text.x  = element_blank(),
        axis.ticks.x= element_blank(),
        axis.title.x = element_text(size=20,face="bold"),
        axis.text.y =  element_text(size=18),
        axis.title.y= element_text(size=18,face="bold"),
        strip.text.x = element_text(size=18))
p
```




## Evaluative Space Grid Categories
The Evaluative Space Grid (ESG) is often divided into four sections: polarizing positive (high positivity, low negativity), polarizing negative (high negativity, low positivity), indifferent (low positivity and negativity), and ambiguous (high positivity and negativity). Here we examine of images belonging to the same ESG categories are placed more closely together. Notably, we see that by around 4-year-of-age children are using ESG category to guide their sorting behaviors.

```{r esg data prep, include=FALSE}

#focus on differences in dist (same v. between categories)
avg_dist_diff_cat_by_subj <- subj_dist_long %>%
  filter(sort!="Practice") %>% 
  group_by(subject, Age, Gender, age_bin,age_group,sort,esg_pair) %>%
  summarize(avg_dist=mean(dist,na.rm=T))  %>%
  ungroup() %>%
  group_by(subject, Age, Gender, age_bin,age_group,sort) %>%
  pivot_wider(names_from=esg_pair,values_from=avg_dist) %>%
  rename(between=diff, within=same) %>%
  mutate(avg_dist_diff=between-within) %>%
  mutate(sort_name=ifelse(sort=="Sort1","Same Individual","Different Individuals")) %>%
  ungroup()
#average across subjects  by age bin
avg_dist_diff_cat_across_subj <- avg_dist_diff_cat_by_subj %>%
  select(-between,-within) %>%
  group_by(age_bin,age_group,sort,sort_name) %>%
  summarize(N=n(),average_dist=mean(avg_dist_diff,na.rm=T),ci_dist=qt(0.975, N-1)*sd(avg_dist_diff,na.rm=T)/sqrt(N))
```

```{r esg plot, warning=FALSE}
p <- ggplot(filter(avg_dist_diff_cat_across_subj, sort!="Practice"),aes(x=sort_name,y=average_dist,fill=age_bin,color=age_bin, linetype=sort_name))+
  geom_errorbar(aes(ymin=average_dist-ci_dist,ymax=average_dist+ci_dist),width=0,size=1)+
  geom_point(size=2.5)+
  facet_wrap(~age_bin,nrow=1, labeller = age_labeller, strip.position = c("bottom"))+
  geom_hline(yintercept=0)+
  theme(legend.position=c(.1,.8), legend.title = element_blank(),
        legend.margin = margin(.5,.5,.5,.5,"cm"),
        legend.background = element_rect(fill="white",size=0.6,
                                         linetype="solid", colour ="black")) +
  scale_color_manual(values=c("#8c510a","#bf812d","#80cdc1","#35978f","#01665e"))+
  guides(color=FALSE) +
  guides(fill=FALSE) +
  ylab("Sorting by ESG Category \nLow<------------------------>High")+
  xlab("Age Group")+
  ylim(-.1,.3) +
  labs(linetype = "Sorting Phase") +
  theme(axis.line = element_blank(),
    axis.text.x  = element_blank(),
        axis.ticks.x= element_blank(),
        axis.title.x = element_text(size=20,face="bold"),
        axis.text.y =  element_text(size=18),
        axis.title.y= element_text(size=18,face="bold"),
        strip.text.x = element_text(size=18)) #+
p
```




We also fit a linear mixed-effects model estimating the average distance between item pairs for children from age (in years, as a continuous predictor; mean-centered), the category match for an image pair (same category pair vs. different category pair; centered), and their interaction. We included a by-participant random intercept and a by-participant random slope for category match. The distances between images belonging to the same vs. different ESG categories increased with age, b = -.03, Wald 95% CI = [-.045, -.02], F(1,103.68) = 43.01, p < .001. As children grow older, they become more likely to sort images belonging to the same emotion category closer together. In follow-up analyses investigating children’s sorting behavior in each of our age groups, we found that 3-year-olds (p = .52) did not reliability sort images from the same ESG category closer together, while 4-year-olds  (b = -.02, Wald 95% CI = [-.04, -.01], F(1,32.745) = 10.55, p < .01), 5-year-olds (b = -.08, Wald 95% CI = [-.011, -.0455], F(1,27) = 23.35, p < .001) and 6-year-olds did (b = -.11, Wald 95% CI = [-.14, -.08], F(1,22) = 56.255, p < .001).


## Mouth Position
Each emotion category had one image with an open mouth and one image with a closed mouth. Since features of the mouth can sometimes influence emotion categorization in children (Caron, Caron & Myers, 1985) we investigated whether mouth position influenced sorting behaviors. Overall we did not find strong evidence that participants were using mouth position to guide their sorting behaviors.
```{r mouth data prep, include=FALSE}
#focus on differences in dist (same v. between categories)
avg_dist_diff_cat_by_subj <- subj_dist_long %>%
  filter(sort!="Practice") %>% 
  group_by(subject, Age, age_bin,age_group,sort,mouth_pair) %>%
  summarize(avg_dist=mean(dist,na.rm=T))  %>%
  ungroup() %>%
  group_by(subject, Age, age_bin,age_group,sort, mouth_pair) %>%
  pivot_wider(names_from=mouth_pair,values_from=avg_dist) %>%
  rename(between=diff, within=same) %>% 
  mutate(avg_dist_diff=between-within) %>%
  mutate(sort_name=ifelse(sort=="Sort1","Same Individual","Different Individuals")) %>%
  ungroup()
#average across subjects  by age bin
avg_dist_diff_cat_across_subj <- avg_dist_diff_cat_by_subj %>%
  select(-between,-within) %>%
  group_by(age_bin,age_group,sort,sort_name) %>%
  summarize(N=n(),average_dist=mean(avg_dist_diff,na.rm=T),ci_dist=qt(0.975, N-1)*sd(avg_dist_diff,na.rm=T)/sqrt(N))
```

```{r mouth plot, warning=FALSE}
# Plot
p <- ggplot(filter(avg_dist_diff_cat_across_subj, sort!="Practice"),aes(x=sort_name,y=average_dist,fill=age_bin,color=age_bin, linetype=sort_name))+
  geom_errorbar(aes(ymin=average_dist-ci_dist,ymax=average_dist+ci_dist),width=0,size=1)+
  geom_point(size=2.5)+
  facet_wrap(~age_bin,nrow=1,labeller = age_labeller, strip.position = c("bottom"))+
  geom_hline(yintercept=0)+
  theme(legend.position=c(.1,.8), legend.title = element_blank(),
        legend.margin = margin(.5,.5,.5,.5,"cm"),
        legend.background = element_rect(fill="white",size=0.6,
                                         linetype="solid", colour ="black")) +
  scale_color_manual(values=c("#8c510a","#bf812d","#80cdc1","#35978f","#01665e"))+
  guides(color=FALSE) +
  guides(fill=FALSE) +
  ylab("Sorting by Mouth Position\nLow<------------------------>High")+
  xlab("Age Group")+
  ylim(-.1,.3) +
  labs(linetype = "Sorting Phase") +
  theme(axis.line = element_blank(),
    axis.text.x  = element_blank(),
        axis.ticks.x= element_blank(),
        axis.title.x = element_text(size=20,face="bold"),
        axis.text.y =  element_text(size=18),
        axis.title.y= element_text(size=18,face="bold"),
        strip.text.x = element_text(size=18)) #+
p

```

## Actor Gender
For actor gender, we only examine the Different Individuals Sort (the Same Individual Sort was all one individual). Here, we look at whether an image was more likely to be placed by another image of the same gender. For instance, participants may have been more likely to place a female face with other female faces.

Overall we don't really find that participants are using actor gender to guide their behaviors. Adults were actually more likely to put actors of the same gender further apart from one another. We believe this is because adults were more likely to use emotion category to guide behavior. During the Different Individuals Sort, each emotion category had one female exemplar and one male exemplar. Thus, using actor gender to guide sorting behavior would cause individuals to use emotion category less.
```{r gender data prep, include=FALSE}
avg_dist_diff_cat_by_subj <- subj_dist_long %>%
  filter(sort=="Sort2") %>% 
  group_by(subject, Age, age_bin,age_group,sort,actor_gender_pair) %>%
  summarize(avg_dist=mean(dist,na.rm=T))  %>%
  ungroup() %>%
  group_by(subject, Age, age_bin,age_group,sort, actor_gender_pair) %>%
  pivot_wider(names_from=actor_gender_pair,values_from=avg_dist) %>%
  rename(between=diff, within=same) %>% 
  mutate(avg_dist_diff=between-within) %>%
  mutate(sort_name="Different Individuals") %>%
  ungroup()
#average across subjects  by age bin
avg_dist_diff_cat_across_subj <- avg_dist_diff_cat_by_subj %>%
  select(-between,-within) %>%
  group_by(age_bin,age_group,sort,sort_name) %>%
  summarize(N=n(),average_dist=mean(avg_dist_diff,na.rm=T),ci_dist=qt(0.975, N-1)*sd(avg_dist_diff,na.rm=T)/sqrt(N))

```

```{r gender plot, warning=FALSE}
p <- ggplot(filter(avg_dist_diff_cat_across_subj, sort!="Practice"),aes(x=sort_name,y=average_dist,fill=age_bin,color=age_bin, linetype=sort_name))+
  geom_errorbar(aes(ymin=average_dist-ci_dist,ymax=average_dist+ci_dist),width=0,size=1)+
  geom_point(size=1.5)+
  facet_wrap(~age_bin,nrow=1, labeller = age_labeller, strip.position = c("bottom"))+
  geom_hline(yintercept=0)+
  theme(legend.position=c(.1,.8), legend.title = element_blank(),
        legend.margin = margin(.5,.5,.5,.5,"cm"),
        legend.background = element_rect(fill="white",size=0.6,
                                         linetype="solid", colour ="black")) +
  scale_color_manual(values=c("#8c510a","#bf812d","#80cdc1","#35978f","#01665e"))+
  guides(color=FALSE) +
  guides(fill=FALSE) +
  ylab("Sorting by Actor Gender \nLow<------------------------>High")+
  xlab("Age Group")+
  ylim(-.1,.3) +
  theme(axis.line = element_blank(),
    axis.text.x  = element_blank(),
        axis.ticks.x= element_blank(),
        axis.title.x = element_text(size=20,face="bold"),
        axis.text.y =  element_text(size=18),
        axis.title.y= element_text(size=18,face="bold"),
        strip.text.x = element_text(size=18)) #+
p

```

# 3. Dimensions of Affect

```{r dim data prep, warning=FALSE}

dimension_data <- subj_dist_long %>% left_join(ratings_pairs)

dimension_data <- dimension_data %>% filter(sort!="Practice") %>% 
  select(subject, sort, Age, age_bin, age_group, dist, items, dist_valence, dist_arousal, dist_pos, dist_neg) %>%
  mutate(subject= as.factor(subject), age_group = as.factor(age_group), age_bin = as.factor(age_bin), sort=as.factor(sort)) %>% 
  mutate(sort2=case_when(
    sort == "Sort1" ~ "Same Individual",
    sort == "Sort2" ~ "Diff. Individual")) %>% 
  mutate(age_bin2=case_when(
    age_bin == '3 to 4' ~ "3 y/o",
    age_bin == '4 to 5' ~ "4 y/o",
    age_bin == '5 to 6' ~ "5 y/o",
    age_bin == '6 to 7' ~ "6 y/o",
    age_bin == 'adults' ~ "adults"))
```

# 3a. Adults versus Children {.tabset} 

Here we display the relationship between the distance between two images based on their image ratings and the average distance that participants placed this images apart. If participants are using the rating of interest, we would expect that images with similar ratings are placed closer together and images with different ratings are placed further apart.

## Valence
```{r}
ggplot(aes(x=dist_valence,y=dist,color=age_group,fill=age_group,linetype=sort2), data=dimension_data) + 
  geom_smooth(method="glm") +
  ylab("Average Distance Between Images")+
  xlab("Diff. in Valence Ratings Between Images")+
  labs(color="Age Group", linetype = "Sort") +
  guides(fill="none") +
  theme(axis.title.x = element_text(size=14,face="bold"),
        axis.text.y =  element_text(size=14),
        axis.title.y= element_text(size=14,face="bold"),
        strip.text.x = element_text(size=14))
```

## Arousal
```{r}
ggplot(aes(x=dist_arousal,y=dist,color=age_group,fill=age_group,linetype=sort2), data=dimension_data) + 
  geom_smooth(method="glm") +
  ylab("Average Distance Between Images")+
  xlab("Diff. in Arousal Ratings Between Images")+
  labs(color="Age Group", linetype = "Sort") +
  guides(fill="none") +
  theme(axis.title.x = element_text(size=14,face="bold"),
        axis.text.y =  element_text(size=14),
        axis.title.y= element_text(size=14,face="bold"),
        strip.text.x = element_text(size=14))
```

## Positivity
```{r}
ggplot(aes(x=dist_pos,y=dist,color=age_group,fill=age_group,linetype=sort2), data=dimension_data) + 
  geom_smooth(method="glm") +
  ylab("Average Distance Between Images")+
  xlab("Diff. in Positivity Ratings Between Images")+
  labs(color="Age Group", linetype = "Sort") +
  guides(fill="none") +
  theme(axis.title.x = element_text(size=14,face="bold"),
        axis.text.y =  element_text(size=14),
        axis.title.y= element_text(size=14,face="bold"),
        strip.text.x = element_text(size=14))
```

## Negativity
```{r}
ggplot(aes(x=dist_neg,y=dist,color=age_group,fill=age_group,linetype=sort2), data=dimension_data) + 
  geom_smooth(method="glm") +
  ylab("Average Distance Between Images")+
  xlab("Diff. in Negativity Ratings Between Images")+
  labs(color="Age Group", linetype = "Sort") +
  guides(fill="none") +
  theme(axis.title.x = element_text(size=14,face="bold"),
        axis.text.y =  element_text(size=14),
        axis.title.y= element_text(size=14,face="bold"),
        strip.text.x = element_text(size=14))
```

# 3b. Children by Age Bin {.tabset} 

Here we display the relationship between the distance between two images based on their image ratings and the average distance that participants placed this images apart. If participants are using the rating of interest, we would expect that images with similar ratings are placed closer together and images with different ratings are placed further apart.

## Valence
```{r}
ggplot(aes(x=dist_valence,y=dist,color=age_bin2,fill=age_bin2), data=filter(dimension_data,age_bin!="adults")) + 
  geom_smooth(method="glm") +
  ylab("Average Distance Between Images")+
  xlab("Diff. in Valence Ratings Between Images")+
  labs(color="Age Group") +
  guides(fill="none") +
  theme(axis.title.x = element_text(size=14,face="bold"),
        axis.text.y =  element_text(size=14),
        axis.title.y= element_text(size=14,face="bold"),
        strip.text.x = element_text(size=14))
```

## Arousal
```{r}
ggplot(aes(x=dist_arousal,y=dist,color=age_bin2,fill=age_bin2), data=filter(dimension_data,age_bin!="adults")) + 
  geom_smooth(method="glm") +
  ylab("Average Distance Between Images")+
  xlab("Diff. in Arousal Ratings Between Images")+
  labs(color="Age Group") +
  guides(fill="none") +
  theme(axis.title.x = element_text(size=14,face="bold"),
        axis.text.y =  element_text(size=14),
        axis.title.y= element_text(size=14,face="bold"),
        strip.text.x = element_text(size=14))
```

## Positivity
```{r}
ggplot(aes(x=dist_pos,y=dist,color=age_bin2,fill=age_bin2), data=filter(dimension_data,age_bin!="adults")) + 
  geom_smooth(method="glm") +
  ylab("Average Distance Between Images")+
  xlab("Diff. in Positivity Ratings Between Images")+
  labs(color="Age Group") +
  guides(fill="none") +
  theme(axis.title.x = element_text(size=14,face="bold"),
        axis.text.y =  element_text(size=14),
        axis.title.y= element_text(size=14,face="bold"),
        strip.text.x = element_text(size=14))
```

## Negativity
```{r}
ggplot(aes(x=dist_neg,y=dist,color=age_bin2,fill=age_bin2), data=filter(dimension_data,age_bin!="adults")) + 
  geom_smooth(method="glm") +
  ylab("Average Distance Between Images")+
  xlab("Diff. in Negativity Ratings Between Images")+
  labs(color="Age Group") +
  guides(fill="none") +
  theme(axis.title.x = element_text(size=14,face="bold"),
        axis.text.y =  element_text(size=14),
        axis.title.y= element_text(size=14,face="bold"),
        strip.text.x = element_text(size=14))
```


# 4. Model Comparisons {.tabset}

```{r}
#load in data
drsq_v_a_ec <- read.csv(here(root_path,"analysis","paper_2020","processed_data","drsqr_v_a_ec.csv"))
drsq_pn_a_ec <- read.csv(here(root_path,"analysis","paper_2020","processed_data","drsqr_pn_a_ec.csv"))

drsq_v_a_ec <- drsq_v_a_ec %>% mutate(age_bin=case_when(
  age_bin == "3 to 4" ~ "3 y/o",
  age_bin == "4 to 5" ~ "4 y/o",
  age_bin == "5 to 6" ~ "5 y/o",
  age_bin == "6 to 7" ~ "6 y/o",
  age_bin == "adults" ~ "adults",
))

drsq_pn_a_ec <- drsq_pn_a_ec %>% mutate(age_bin=case_when(
  age_bin == "3 to 4" ~ "3 y/o",
  age_bin == "4 to 5" ~ "4 y/o",
  age_bin == "5 to 6" ~ "5 y/o",
  age_bin == "6 to 7" ~ "6 y/o",
  age_bin == "adults" ~ "adults",
))

```

## Valence plots
The left plot displays the full Model R-squared for each age group. The right plot displays the Delta R-squared for each predictor of sorting behavior. Error bars represent bootstrapped 95% confidence intervals.
```{r}
pA <- ggplot(drsq_v_a_ec,aes(minus_model_name,delta_rsq,  fill=minus_model_name))+
  geom_bar(stat="identity",color="black")+
  geom_hline(yintercept=0,color="black")+
  geom_errorbar(aes(ymin=delta_rsq_lower,ymax=delta_rsq_upper),width=0.1)+
  scale_fill_viridis_d(
                    name = " ", #Model Predictor
                    limits=c("dist_valence","dist_arousal","emotion_pair_same"),
                    breaks=c("dist_valence","dist_arousal","emotion_pair_same"),
                    labels=c("Valence","Arousal","Emotion Category"))+
  scale_x_discrete(
    limits=c("dist_valence","dist_arousal","emotion_pair_same"),
    breaks=c("dist_valence","dist_arousal","emotion_pair_same"),
    labels=c("Valence","Arousal","Same Emotion Category"))+
  theme(axis.text.x  = element_text(angle=90, vjust=0.5,size=18),
        axis.title.x = element_text(size=14,face="bold"),
        axis.text.y =  element_text(size=14),
        axis.title.y= element_text(size=14,face="bold"),
        strip.text.x = element_text(size=14),
        legend.text =element_text(size=14),
        legend.title =element_text(size=16,face="bold"))+
  theme(axis.title.x=element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank())+
  ylab("Delta R Squared")+
  facet_wrap(~age_bin,nrow=1, strip.position = "top")+
  theme(legend.position=c(0.1,0.8))

pB <- ggplot(unique(select(drsq_v_a_ec,age_bin,full_model_rsq)),aes(age_bin,full_model_rsq,  fill=age_bin))+
  geom_bar(stat="identity",color="black")+
  geom_hline(yintercept=0,color="black")+
  theme(axis.text.x  = element_text(angle=90, vjust=0.5,size=14),
        axis.title.x = element_text(size=14,face="bold"),
        axis.text.y =  element_text(size=14),
        axis.title.y= element_text(size=14,face="bold"),
        strip.text.x = element_text(size=14))+
  theme(axis.title.x=element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank())+
  ylab("Model R Squared")+
  xlab("Age Group")+
theme(legend.position=c(0.1,0.8))
plot_grid(pB,pA)

```

## Pos-Neg plots
The left plot displays the full Model R-squared for each age group. The right plot displays the Delta R-squared for each predictor of sorting behavior. Error bars represent bootstrapped 95% confidence intervals.
```{r}
pA <- ggplot(drsq_pn_a_ec,aes(minus_model_name,delta_rsq,  fill=minus_model_name))+
  geom_bar(stat="identity",color="black")+
  geom_hline(yintercept=0,color="black")+
  geom_errorbar(aes(ymin=delta_rsq_lower,ymax=delta_rsq_upper),width=0.1)+
  scale_fill_brewer(palette="Set1",
                    name = " ", #Model Predictor
                    limits=c("dist_pos","dist_neg","dist_arousal","emotion_pair_same"),
                    breaks=c("dist_pos","dist_neg","dist_arousal","emotion_pair_same"),
                    labels=c("Positive","Negative","Arousal","Same Emotion Category"))+
  scale_x_discrete(
    limits=c("dist_pos","dist_neg","dist_arousal","emotion_pair_same"),
    breaks=c("dist_pos","dist_neg","dist_arousal","emotion_pair_same"),
    labels=c("Positive","Negative","Arousal","Same Emotion Category"))+
  theme(axis.text.x  = element_text(angle=90, vjust=0.5,size=14),
        axis.title.x = element_text(size=14,face="bold"),
        axis.text.y =  element_text(size=14),
        axis.title.y= element_text(size=14,face="bold"),
        strip.text.x = element_text(size=14))+
  theme(axis.title.x=element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank())+
  ylab("Delta R Squared")+
  facet_wrap(~age_bin,nrow=1)+
  theme(legend.position=c(0.1,0.9))

pB <- ggplot(unique(select(drsq_pn_a_ec,age_bin,full_model_rsq)),aes(age_bin,full_model_rsq,  fill=age_bin))+
  geom_bar(stat="identity",color="black")+
  geom_hline(yintercept=0,color="black")+
  theme(axis.text.x  = element_text(angle=90, vjust=0.5,size=18),
        axis.title.x = element_text(size=14,face="bold"),
        axis.text.y =  element_text(size=14),
        axis.title.y= element_text(size=14,face="bold"),
        strip.text.x = element_text(size=14))+
  theme(axis.title.x=element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank())+
  ylab("Model R Squared")+
  xlab("Age Group")+
  theme(legend.position=c(0.1,0.8))
plot_grid(pB,pA)
```



# 5a. Multidimensional Scaling: Same Individual Sort {.tabset} 
Here we present the same multidimensionsal scaling (MDS) results as in the manuscript. However, we provide additional visualizations. Reviewers can now see both faces and emotion category labels, as well as which cluster each face and image were a part of (from the hierarchical clustering k=3 solution)
```{r mds data prep, include=FALSE, warning=FALSE}
#average across all distances
avg_dist_long <- subj_dist_long %>%
  group_by(sort,age_group,item1,item2) %>%
  summarize(avg_dist=mean(dist)) %>%
  ungroup() %>%
  mutate(sort=as.character(sort),age_group=as.character(age_group),item1=as.character(item1),item2=as.character(item2))

#average distance object organized by sorting group
avg_dist <- avg_dist_long %>%
  group_by(sort,age_group) %>%
  nest() %>%
  mutate(dist_obj = map(data, long_to_dist))

#For Scree Plots
cur_dist <- filter(avg_dist,sort=="Sort1"&age_group=="kids")$dist_obj[[1]]
mds1 <- cmdscale(cur_dist, eig=TRUE)

cur_dist <- filter(avg_dist,sort=="Sort1"&age_group=="adults")$dist_obj[[1]]
mds2 <- cmdscale(cur_dist, eig=TRUE)

cur_dist <- filter(avg_dist,sort=="Sort2"&age_group=="kids")$dist_obj[[1]]
mds3 <- cmdscale(cur_dist, eig=TRUE)

cur_dist <- filter(avg_dist,sort=="Sort2"&age_group=="adults")$dist_obj[[1]]
mds4 <- cmdscale(cur_dist, eig=TRUE)

```


```{r mds plot prep}
cur_dist <- filter(avg_dist,sort=="Sort1"&age_group=="kids")$dist_obj[[1]]
sort1_kids_cmd <- data.frame(cmdscale(cur_dist, k=2))

# for image rating vectors
ratings1 <- ratings %>% filter(sort=="Sort1") %>%  select(image, pos, neg, valence, arousal) %>% rename(positivity=pos, negativity=neg)
sort1k_fit <- envfit(sort1_kids_cmd,ratings1, permutations = 1000)

# prepare images for plot
cur_cluster <- hclust(cur_dist)
image_paths <- paste(here("experiment","stimuli_sort1"),"/",labels(cur_cluster),".png",sep="")
cur_images <- data.frame(
  label=labels(cur_cluster),
  image=image_paths,
  x=seq(1,length(image_paths)),
  y=rep(c(-0.05,-0.1),length(image_paths)/2))
sort1_images <- cur_images
cur_cmd <- data.frame(cmdscale(cur_dist))
cur_cmd$label <- rownames(cur_cmd)
cur_cmd <- merge(cur_cmd,sort1_images)

#include correlations at correct angles
data.scores = as.data.frame(scores(sort1_kids_cmd))
en_coord_cont = as.data.frame(scores(sort1k_fit, "vectors")) * ordiArrowMul(sort1k_fit)

#Plot MDS
p1 <- ggplot(data=cur_cmd,aes(X1,X2))+
  geom_image(data=cur_cmd, aes(image=image), size=0.08)+
  ylab("Dimension 2")+
  xlab("Dimension 1")+
  geom_segment(data = en_coord_cont,aes(x = 0, y = 0, xend = X1/3, yend = X2/3, color=row.names(en_coord_cont)), 
               size =1, alpha = 0.8, arrow = arrow(length = unit(0.03, "npc"))) + 
  geom_text(data = en_coord_cont, aes(x = X1/3, y = (X2/3) - .01), colour = "black",
            fontface = "bold", label = row.names(en_coord_cont))+
  theme(axis.text.x  = element_text(angle=90, vjust=0.5,size=18),
        axis.title.x = element_text(size=18,face="bold"),
        axis.text.y =  element_text(size=16),
        axis.title.y= element_text(size=18,face="bold"),
        strip.text.x = element_text(size=16))+
  ggtitle("Children")+
  theme(legend.position="none", plot.title = element_text(hjust = 0.5,size=18))

p1b <- ggplot(data=cur_cmd,aes(X1,X2))+
  geom_image(data=cur_cmd, aes(image=image), size=0.08)+
  ylab("Dimension 2")+
  xlab("Dimension 1")+
  annotate("rect", xmin=-.25,xmax=-.02, ymin=-.2,ymax=.2,alpha=.15,fill='red')+
  annotate("rect", xmin= .13,xmax=0.22, ymin=-.13,ymax=0,alpha=.15,fill='blue')+
  annotate("rect", xmin=-.02,xmax=0.13, ymin=-.05,ymax=.2,alpha=.15,fill='green')+
  theme(axis.text.x  = element_text(angle=90, vjust=0.5,size=18),
        axis.title.x = element_text(size=18,face="bold"),
        axis.text.y =  element_text(size=16),
        axis.title.y= element_text(size=18,face="bold"),
        strip.text.x = element_text(size=16))+
  ggtitle("Children")+
  theme(legend.position="none", plot.title = element_text(hjust = 0.5,size=18))

#adults
cur_dist <- filter(avg_dist,sort=="Sort1"&age_group=="adults")$dist_obj[[1]]
sort1_adults_cmd <- data.frame(cmdscale(cur_dist, k=2))

# for image rating vectors
sort1a_fit <- envfit(sort1_adults_cmd,ratings1, permutations=1000)
data.scores = as.data.frame(scores(sort1_adults_cmd))
en_coord_cont = as.data.frame(scores(sort1a_fit, "vectors")) * ordiArrowMul(sort1a_fit)

#prepare images
cur_cmd <- data.frame(cmdscale(cur_dist))
cur_cmd$label <- rownames(cur_cmd)
cur_cmd <- merge(cur_cmd,sort1_images)

#Plot MDS
p2 <- ggplot(data=cur_cmd,aes(X1,X2))+
  geom_image(data=cur_cmd, aes(image=image), size=0.08)+
  ylab("Dimension 2")+
  xlab("Dimension 1")+
  geom_segment(data = en_coord_cont,aes(x = 0, y = 0, xend = X1/3, yend = X2/3, color=row.names(en_coord_cont)), 
               size =1, alpha = 0.8, arrow = arrow(length = unit(0.03, "npc"))) + #ends="both" for arrow
  geom_text(data = en_coord_cont, aes(x = X1/3, y = (X2/3) - .01), colour = "black",
            fontface = "bold", label = row.names(en_coord_cont))+
  theme(axis.text.x  = element_text(angle=90, vjust=0.5,size=18),
        axis.title.x = element_text(size=18,face="bold"),
        axis.text.y =  element_text(size=16),
        axis.title.y= element_text(size=18,face="bold"),
        strip.text.x = element_text(size=16))+
  ggtitle("Adults")+
  theme(legend.position="none", plot.title = element_text(hjust = 0.5,size=18))

p2b <- ggplot(data=cur_cmd,aes(X1,X2))+
  geom_image(data=cur_cmd, aes(image=image), size=0.08)+
  ylab("Dimension 2")+
  xlab("Dimension 1")+
  annotate("rect", xmin=-.35,xmax=-.14, ymin=-.07,ymax=.2,alpha=.15,fill='red')+
  annotate("rect", xmin=.15,xmax=0.4, ymin=-.07,ymax=.2,alpha=.15,fill='blue')+
  annotate("rect", xmin=-.2,xmax=0.18, ymin=-.25,ymax=-.07,alpha=.15,fill='green')+
  theme(axis.text.x  = element_text(angle=90, vjust=0.5,size=18),
        axis.title.x = element_text(size=18,face="bold"),
        axis.text.y =  element_text(size=16),
        axis.title.y= element_text(size=18,face="bold"),
        strip.text.x = element_text(size=16))+
  ggtitle("Adults")+
  theme(legend.position="none", plot.title = element_text(hjust = 0.5,size=18))
```

```{r label prep}
library(ggrepel)

#Children
cur_dist <- filter(avg_dist,sort=="Sort1"&age_group=="kids")$dist_obj[[1]]
sort1_kids_cmd <- data.frame(cmdscale(cur_dist, k=2))
cur_cmd <- data.frame(cmdscale(cur_dist))
cur_cmd$label <- rownames(cur_cmd)
cur_cmd <- merge(cur_cmd,sort1_images)

s1k <- hclust(cur_dist, method = "ward.D2") #get hclust solution
s1k <- as.data.frame(cutree(s1k, 3))
s1k <- s1k %>% tibble::rownames_to_column("label") %>% rename(cluster3 = `cutree(s1k, 3)`)
s1k <- s1k %>% mutate(clean_label =
  str_replace_all(label, c("M07"="","_o"=" 1", "_c"=" 2", "ang"="angry","calm"="calm","disg"="disgust",
                    "exc"="excited","fear"="fear", "hap"="happy",
                    "neut"="neutral","sad"="sad", "surp"="surprise")))

cur_cmd <- cur_cmd %>%  #combine hclust and mds
  dplyr:::inner_join(s1k, by = c("label"))

p1c <- ggplot(data=cur_cmd,aes(X1,X2,label=clean_label,color=as.factor(cluster3)))+
  geom_text_repel(fontface = "bold", min.segment.length = .3)+
  ylab("Dimension 2")+
  xlab("Dimension 1")+
  theme(axis.text.x  = element_text(angle=90, vjust=0.5,size=18),
        axis.title.x = element_text(size=18,face="bold"),
        axis.text.y =  element_text(size=16),
        axis.title.y= element_text(size=18,face="bold"),
        strip.text.x = element_text(size=16))+
  ggtitle("Children")+
  theme(legend.position="none", plot.title = element_text(hjust = 0.5,size=18))

#Adults
cur_dist <- filter(avg_dist,sort=="Sort1"&age_group=="adults")$dist_obj[[1]]
sort1_adults_cmd <- data.frame(cmdscale(cur_dist, k=2))
cur_cmd <- data.frame(cmdscale(cur_dist))
cur_cmd$label <- rownames(cur_cmd)
cur_cmd <- merge(cur_cmd,sort1_images)

s1a <- hclust(cur_dist, method = "ward.D2") #get hclust solution
s1a <- as.data.frame(cutree(s1a, 3))
s1a <- s1a %>% tibble::rownames_to_column("label") %>% rename(cluster3 = `cutree(s1a, 3)`)
s1a <- s1a %>% mutate(clean_label =
  str_replace_all(label, c("M07"="","_o"=" 1", "_c"=" 2", "ang"="angry","calm"="calm","disg"="disgust",
                    "exc"="excited","fear"="fear", "hap"="happy",
                    "neut"="neutral","sad"="sad", "surp"="surprise")))

cur_cmd <- cur_cmd %>%  #combine hclust and mds
  dplyr:::inner_join(s1a, by = c("label"))

#changing for color reasons only
cur_cmd <- cur_cmd %>% mutate(cluster3=case_when(
  cluster3 == 1~1,
  cluster3 == 2~3,
  cluster3 == 3~2))

p2c <- ggplot(data=cur_cmd,aes(X1,X2,label=clean_label,color=as.factor(cluster3)))+
  geom_text_repel(fontface = "bold", min.segment.length = .3)+
  ylab("Dimension 2")+
  xlab("Dimension 1")+
  theme(axis.text.x  = element_text(angle=90, vjust=0.5,size=18),
        axis.title.x = element_text(size=18,face="bold"),
        axis.text.y =  element_text(size=16),
        axis.title.y= element_text(size=18,face="bold"),
        strip.text.x = element_text(size=16))+
  ggtitle("Adults")+
  theme(legend.position="none", plot.title = element_text(hjust = 0.5,size=18))

```

## Faces with image ratings
```{r}
plot_grid(p1,p2,labels=c("A","B"),label_size=20)

```

## Faces with Hierarchical Clustering solution
```{r}
plot_grid(p1b,p2b,labels=c("A","B"),label_size=20)


```


Colors highlight k=3 hierarchical clustering solutions for each age group. We used the three colors to highlight commonalities among the different hierarchical clustering solutions. For instance, red clusters contain anger and disgust images, green clusters contain certain fear and neutral images, and blue images contain certain happy, calm, and surprise images. However, there was nothing statistically to indicate “this is the red cluster” across all dendrograms.

## Labels with Hierarchical Clustering solution
```{r label plot}
plot_grid(p1c,p2c,labels=c("A","B"),label_size=20)


```


Colors highlight the three cluster solutions for each age-bin. We used the three colors to highlight commonalities among the different hierarchical clustering solutions. For instance, red clusters contain anger and disgust images, green clusters contain certain fear and neutral images, and blue images contain certain happy, calm, and surprise images. However, there was nothing statistically to indicate “this is the red cluster” across all dendrograms. The numbers 1 and 2 indicate that the images had open and closed mouths, respectively. Labels are slightly jittered for readability (indicated by lines).

# 5b. Multidimensional Scaling: Different Individuals Sort {.tabset} 
Here we present the same multidimensionsal scaling (MDS) results as in the manuscript. However, we provide additional visualizations. Reviewers can now see both faces and emotion category labels, as well as which cluster each face and image were a part of (from the hierarchical clustering k=3 solution)
```{r mds plot prep2}
#### Children ####
cur_dist <- filter(avg_dist,sort=="Sort2"&age_group=="kids")$dist_obj[[1]]
sort2_kids_cmd <- data.frame(cmdscale(cur_dist, k=2))

# get vectors of image ratings
ratings2 <- ratings %>% filter(sort=="Sort2") %>%  select(image, pos, neg, valence, arousal) %>% rename(positivity=pos, negativity=neg)
sort2k_fit <- envfit(sort2_kids_cmd,ratings2, permutations=1000)
data.scores = as.data.frame(scores(sort2_kids_cmd))
en_coord_cont = as.data.frame(scores(sort2k_fit, "vectors")) * ordiArrowMul(sort2k_fit)

#create image paths
cur_cluster <- hclust(cur_dist)
image_paths <- paste(here("experiment","stimuli_sort2"),"/",labels(cur_cluster),".png",sep="")
cur_images <- data.frame(
  label=labels(cur_cluster),
  image=image_paths,
  x=seq(1,length(image_paths)),
  y=rep(c(-0.05,-0.1),length(image_paths)/2))
sort2_images <- cur_images
cur_cmd <- data.frame(cmdscale(cur_dist))
cur_cmd$label <- rownames(cur_cmd)
cur_cmd <- merge(cur_cmd,sort2_images)

#Plot MDS
p1 <- ggplot(data=cur_cmd,aes(X1,X2))+
  geom_image(data=cur_cmd, aes(image=image), size=0.08)+
  ylab("Dimension 2")+
  xlab("Dimension 1")+
  geom_segment(data = en_coord_cont,aes(x = 0, y = 0, xend = X1/3, yend = X2/3, color=row.names(en_coord_cont)), 
               size =1, alpha = 0.8, arrow = arrow(length = unit(0.03, "npc"))) + #ends="both" for arrow
  geom_text(data = en_coord_cont, aes(x = X1/3, y = (X2/3) - .01), colour = "black",
            fontface = "bold", label = row.names(en_coord_cont))+
  theme(axis.text.x  = element_text(angle=90, vjust=0.5,size=18),
        axis.title.x = element_text(size=18,face="bold"),
        axis.text.y =  element_text(size=16),
        axis.title.y= element_text(size=18,face="bold"),
        strip.text.x = element_text(size=16))+
  ggtitle("Children")+
  theme(legend.position="none", plot.title = element_text(hjust = 0.5,size=18))

p1b <- ggplot(data=cur_cmd,aes(X1,X2))+
  geom_image(data=cur_cmd, aes(image=image), size=0.08)+
  ylab("Dimension 2")+
  xlab("Dimension 1")+
  annotate("rect", xmin=-.26,xmax=-.12, ymin=-.18,ymax=.12,alpha=.15,fill='red')+
  annotate("rect", xmin=0.05,xmax=0.2, ymin=-.2,ymax=0.05,alpha=.15,fill='blue')+
  annotate("rect", xmin=-.05,xmax=0.12, ymin=0.05,ymax=.2,alpha=.15,fill='green')+
  theme(axis.text.x  = element_text(angle=90, vjust=0.5,size=18),
        axis.title.x = element_text(size=18,face="bold"),
        axis.text.y =  element_text(size=16),
        axis.title.y= element_text(size=18,face="bold"),
        strip.text.x = element_text(size=16))+
  ggtitle("Children")+
  theme(legend.position="none", plot.title = element_text(hjust = 0.5,size=18))


#### Adults ####
cur_dist <- filter(avg_dist,sort=="Sort2"&age_group=="adults")$dist_obj[[1]]
sort2_adults_cmd <- data.frame(cmdscale(cur_dist, k=2))

# for image rating vectors
sort2a_fit <- envfit(sort2_adults_cmd,ratings2, permutations=1000)
data.scores = as.data.frame(scores(sort2_adults_cmd))
en_coord_cont = as.data.frame(scores(sort2a_fit, "vectors")) * ordiArrowMul(sort2a_fit)

# prepare images
cur_cmd <- data.frame(cmdscale(cur_dist))
cur_cmd$label <- rownames(cur_cmd)
cur_cmd <- merge(cur_cmd,sort2_images)

#Plot MDS
p2 <- ggplot(data=cur_cmd,aes(X1,X2))+
  geom_image(data=cur_cmd, aes(image=image), size=0.08)+
  ylab("Dimension 2")+
  xlab("Dimension 1")+
  geom_segment(data = en_coord_cont,aes(x = 0, y = 0, xend = X1/3, yend = X2/3, color=row.names(en_coord_cont)), 
               size =1, alpha = 0.8, arrow = arrow(length = unit(0.03, "npc"))) + #ends="both" for arrow
  geom_text(data = en_coord_cont, aes(x = X1/3, y = (X2/3) - .02), colour = "black",
            fontface = "bold", label = row.names(en_coord_cont))+
  theme(axis.text.x  = element_text(angle=90, vjust=0.5,size=18),
        axis.title.x = element_text(size=18,face="bold"),
        axis.text.y =  element_text(size=16),
        axis.title.y= element_text(size=18,face="bold"),
        strip.text.x = element_text(size=16))+
  ggtitle("Adults")+
  theme(legend.position="none", plot.title = element_text(hjust = 0.5,size=18))

p2b <- ggplot(data=cur_cmd,aes(X1,X2))+
  geom_image(data=cur_cmd, aes(image=image), size=0.08)+
  ylab("Dimension 2")+
  xlab("Dimension 1")+
  annotate("rect", xmin=-.4,xmax=-.27, ymin=-.13,ymax=0.02,alpha=.15,fill='blue')+
  annotate("rect", xmin=.19,xmax=0.35, ymin=-.25,ymax=-.12,alpha=.15,fill='red')+
  annotate("rect", xmin=-.17,xmax=0.20, ymin=-0.11,ymax=.31,alpha=.15,fill='green')+
  theme(axis.text.x  = element_text(angle=90, vjust=0.5,size=18),
        axis.title.x = element_text(size=18,face="bold"),
        axis.text.y =  element_text(size=16),
        axis.title.y= element_text(size=18,face="bold"),
        strip.text.x = element_text(size=16))+
  ggtitle("Adults")+
  theme(legend.position="none", plot.title = element_text(hjust = 0.5,size=18))


```

```{r}
library(ggrepel)

#Children
cur_dist <- filter(avg_dist,sort=="Sort2"&age_group=="kids")$dist_obj[[1]]
sort2_kids_cmd <- data.frame(cmdscale(cur_dist, k=2))
cur_cmd <- data.frame(cmdscale(cur_dist))
cur_cmd$label <- rownames(cur_cmd)
cur_cmd <- merge(cur_cmd,sort2_images)

s2k <- hclust(cur_dist, method = "ward.D2") #get hclust solution
s2k <- as.data.frame(cutree(s2k, 3))
s2k <- s2k %>% tibble::rownames_to_column("label") %>% rename(cluster3 = `cutree(s2k, 3)`)
s2k <- s2k %>% mutate(clean_label =
  str_replace_all(label, c("F01"="","F04"="","F07"="","F10"="", "F13"="",
                    "F14"="","F15"="", "F17"="","F22"="","M02"="","M04"="",
                    "M03"="","M05"="","M08"="","M12"="","M14"="","M15"="",
                    "M17"="","_o"=" 1", "_c"=" 2", "ang"="angry", "calm"="calm",
                    "disg"="disgust", "exc"="excited", "fear"="fear",
                    "hap"="happy", "neut"="neutral", "sad"="sad",
                    "surp"="surprise")))

cur_cmd <- cur_cmd %>% #combine hclust and mds
  dplyr:::inner_join(s2k, by = c("label"))

#changing for color reasons only
cur_cmd <- cur_cmd %>% mutate(cluster3=case_when(
  cluster3 == 1~3,
  cluster3 == 2~1,
  cluster3 == 3~2))

p1c <- ggplot(data=cur_cmd,aes(X1,X2,label=clean_label,color=as.factor(cluster3)))+
  geom_text_repel(fontface = "bold", min.segment.length = .3)+
  ylab("Dimension 2")+
  xlab("Dimension 1")+
  theme(axis.text.x  = element_text(angle=90, vjust=0.5,size=18),
        axis.title.x = element_text(size=18,face="bold"),
        axis.text.y =  element_text(size=16),
        axis.title.y= element_text(size=18,face="bold"),
        strip.text.x = element_text(size=16))+
  ggtitle("Children")+
  theme(legend.position="none", plot.title = element_text(hjust = 0.5,size=18))

#Adults
cur_dist <- filter(avg_dist,sort=="Sort2"&age_group=="adults")$dist_obj[[1]]
sort2_adults_cmd <- data.frame(cmdscale(cur_dist, k=2))
cur_cmd <- data.frame(cmdscale(cur_dist))
cur_cmd$label <- rownames(cur_cmd)
cur_cmd <- merge(cur_cmd,sort2_images)

s2a <- hclust(cur_dist, method = "ward.D2") #get hclust solution
s2a <- as.data.frame(cutree(s2a, 3))
s2a <- s2a %>% tibble::rownames_to_column("label") %>% rename(cluster3 = `cutree(s2a, 3)`)
s2a <- s2a %>% mutate(clean_label =
  str_replace_all(label, c("F01"="","F04"="","F07"="","F10"="", "F13"="",
                    "F14"="","F15"="", "F17"="","F22"="","M02"="","M04"="",
                    "M03"="","M05"="","M08"="","M12"="","M14"="","M15"="",
                    "M17"="","_o"=" 1", "_c"=" 2", "ang"="angry", "calm"="calm",
                    "disg"="disgust", "exc"="excited", "fear"="fear",
                    "hap"="happy", "neut"="neutral", "sad"="sad",
                    "surp"="surprise")))

cur_cmd <- cur_cmd %>% #combine hclust and mds
  dplyr:::inner_join(s2a, by = c("label"))

#changing for color reasons only
cur_cmd <- cur_cmd %>% mutate(cluster3=case_when(
  cluster3 == 1~2,
  cluster3 == 2~1,
  cluster3 == 3~3))

p2c <- ggplot(data=cur_cmd,aes(X1,X2,label=clean_label,color=as.factor(cluster3)))+
  geom_text_repel(fontface = "bold", min.segment.length = .3)+
  ylab("Dimension 2")+
  xlab("Dimension 1")+
  theme(axis.text.x  = element_text(angle=90, vjust=0.5,size=18),
        axis.title.x = element_text(size=18,face="bold"),
        axis.text.y =  element_text(size=16),
        axis.title.y= element_text(size=18,face="bold"),
        strip.text.x = element_text(size=16))+
  ggtitle("Adults")+
  theme(legend.position="none", plot.title = element_text(hjust = 0.5,size=18))
```

## Faces with image ratings
```{r}
plot_grid(p1,p2,labels=c("C","D"),label_size=20)


```

## Faces with Hierarchical Clustering solution

```{r}
plot_grid(p1b,p2b,labels=c("C","D"),label_size=20)


```


Colors highlight k=3 hierarchical clustering solutions for each age group. We used the three colors to highlight commonalities among the different hierarchical clustering solutions. For instance, red clusters contain anger and disgust images, green clusters contain certain fear and neutral images, and blue images contain certain happy, calm, and surprise images. However, there was nothing statistically to indicate “this is the red cluster” across all dendrograms.


## Labels with Hierarchical Clustering solution
 
```{r label plot2}
plot_grid(p1c,p2c,labels=c("A","B"),label_size=20)


```


Colors highlight the three cluster solutions for each age-bin. We used the three colors to highlight commonalities among the different hierarchical clustering solutions. For instance, red clusters contain anger and disgust images, green clusters contain certain fear and neutral images, and blue images contain certain happy, calm, and surprise images. However, there was nothing statistically to indicate “this is the red cluster” across all dendrograms. The numbers 1 and 2 indicate that the images had open and closed mouths, respectively. Labels are slightly jittered for readability (indicated by lines).

# 5c. Multidimensional Scaling: How many dimensions? {.tabset}
We found that the 2-dimensional MDS solution had the best fit. We also felt that the 2-dimensional structure was most appropriate given the nature of the task (e.g., sorting all images together in a 2-dimensional space)
```{r eig data prep}
m1 <- data.frame(mds1$eig)
m2 <- data.frame(mds2$eig)
m3 <- data.frame(mds3$eig)
m4 <- data.frame(mds4$eig)

m1 <- m1 %>% rownames_to_column() %>% rename(num_dim=rowname, eig=mds1.eig) %>% mutate(sort="Same Individual",age="Children")
m2 <- m2 %>% rownames_to_column() %>% rename(num_dim=rowname, eig=mds2.eig) %>% mutate(sort="Same Individual",age="Adults")
m3 <- m3 %>% rownames_to_column() %>% rename(num_dim=rowname, eig=mds3.eig) %>% mutate(sort="Different Individuals",age="Children")
m4 <- m4 %>% rownames_to_column() %>% rename(num_dim=rowname, eig=mds4.eig) %>% mutate(sort="Different Individuals",age="Adults")
m_full <- rbind(m1,m2,m3,m4)
m_full$num_dim=as.numeric(m_full$num_dim)
```

```{r plot eig,}
p <- ggplot(data=m_full,aes(x=num_dim,y=eig,fill=age, color=age)) + 
  geom_area() +
  facet_wrap(~sort) +
  geom_vline(xintercept=2)+
  ylab("Eigenvalue")+
  xlab("Number of Dimensions") +
  scale_x_continuous(breaks=c(1,2,3,4,5,6,7,8,9,11,13,15,17))
 p 

```




# 6a. Hierarchical Clustering: Additional Dendrograms {.tabset}
We present additional dendrograms comparison adults and children (collapsing across all age groups). A dendrogram is a way of visualizing a hierarchical clustering solution. It displays all clustering solutions - in our case, from k =2 to k =18. Each face participants sorted is a “node” on the dendrogram and when it merges with another face (represented by a horizontal line) another “node” is formed. Clusters are determined by the vertical height of the branches in a dendrogram, not by which labels are closest to one another laterally. Thus, faces that were found to be the most similar would be connected as a node with a very low height. The k = 3 cluster that we display shows what happens if a horizontal line was drawn that only passed through 3 vertical lines.

Colors highlight the three cluster solutions for each age-bin. We used the three colors to highlight commonalities among the different dendrograms. For instance, red clusters contain anger and disgust images, green clusters contain certain fear and neutral images, and blue images contain certain happy, calm, and surprise images. However, there was nothing statistically to indicate “this is the red cluster” across all dendrograms. The numbers 1 and 2 indicate that the images had open and closed mouths, respectively.

```{r data prep hclust, warning=FALSE}
#average across all distances
avg_dist_long_byGroup <- subj_dist_long %>%
  group_by(sort,age_group,item1,item2) %>%
  summarize(avg_dist=mean(dist)) %>%
  ungroup() %>%
  mutate(sort=as.character(sort),age_group=as.character(age_group),item1=as.character(item1),item2=as.character(item2))

#average distance object organized by sorting group
avg_dist_byGroup <- avg_dist_long_byGroup %>%
  group_by(sort,age_group) %>%
  nest() %>%
  mutate(dist_obj = map(data, long_to_dist))

#average across all distances
avg_dist_long_byBin <- subj_dist_long %>%
  group_by(sort,age_bin,item1,item2) %>%
  summarize(avg_dist=mean(dist)) %>%
  ungroup() %>%
  mutate(sort=as.character(sort),age_bin=as.character(age_bin),item1=as.character(item1),item2=as.character(item2))

#average distance object organized by sorting group
avg_dist_byBin <- avg_dist_long_byBin %>%
  group_by(sort,age_bin) %>%
  nest() %>%
  mutate(dist_obj = map(data, long_to_dist))
```


## Same Individual Sort: Adults and Children

```{r}
#distance matrix objects
s1k_dist <- avg_dist_byGroup %>%
  dplyr::filter(sort=="Sort1"&age_group=="kids") %>%
  ungroup() %>%
  select(dist_obj)
s1k_dist <- s1k_dist$dist_obj[[1]]

s1a_dist <- avg_dist_byGroup %>%
  dplyr::filter(sort=="Sort1"&age_group=="adults") %>%
  ungroup() %>%
  select(dist_obj)
s1a_dist <- s1a_dist$dist_obj[[1]]

# hclust solution
s1k <- s1k_dist %>%
  hclust(method = "ward.D2")

s1a <- s1a_dist %>%
  hclust(method = "ward.D2")

#clean up labels
s1k[["labels"]] <- s1k[["labels"]] %>%
  str_replace_all(c("M07"="","_o"=" 1", "_c"=" 2", "ang"="angry","calm"="calm","disg"="disgust",
                    "exc"="excited","fear"="fear", "hap"="happy",
                    "neut"="neutral","sad"="sad", "surp"="surprise"))

s1a[["labels"]] <- s1a[["labels"]] %>%
  str_replace_all(c("M07"="","_o"=" 1", "_c"=" 2", "ang"="angry","calm"="calm","disg"="disgust",
                    "exc"="excited","fear"="fear", "hap"="happy",
                    "neut"="neutral","sad"="sad", "surp"="surprise"))

# dendrograms
s1k_dend <- s1k %>% as.dendrogram
s1a_dend <- s1a %>% as.dendrogram

# plots
par(mfrow = c(1,2))
s1a_dend %>% set("branches_lwd", 2) %>%  set("branches_k_color",value=c(4,3,2), k=3) %>% plot(main="Adults")
s1k_dend %>% set("branches_lwd", 2) %>%  set("branches_k_color",value=c(2,4,3), k=3)%>% plot(main="Children")
```

## Different Individuals Sort: Adults and Children
```{r}
s2k_dist <- avg_dist_byGroup %>%
  dplyr::filter(sort=="Sort2"&age_group=="kids") %>%
  ungroup() %>%
  select(dist_obj)
s2k_dist <- s2k_dist$dist_obj[[1]]

s2a_dist <- avg_dist_byGroup %>%
  dplyr::filter(sort=="Sort2"&age_group=="adults") %>%
  ungroup() %>%
  select(dist_obj)
s2a_dist <- s2a_dist$dist_obj[[1]]

#hclust objects
s2k <- s2k_dist %>%
  hclust(method = "ward.D2")

s2a <- s2a_dist %>%
  hclust(method = "ward.D2")


#clean up labels
s2k[["labels"]] <- s2k[["labels"]] %>%
  str_replace_all(c("F01"="","F04"="","F07"="","F10"="", "F13"="",
                    "F14"="","F15"="", "F17"="","F22"="","M02"="","M04"="",
                    "M03"="","M05"="","M08"="","M12"="","M14"="","M15"="",
                    "M17"="","_o"=" 1", "_c"=" 2", "ang"="angry", "calm"="calm",
                    "disg"="disgust", "exc"="excited", "fear"="fear",
                    "hap"="happy", "neut"="neutral", "sad"="sad",
                    "surp"="surprise"))

s2a[["labels"]] <- s2a[["labels"]] %>%
  str_replace_all(c("F01"="","F04"="","F07"="","F10"="", "F13"="",
                    "F14"="","F15"="", "F17"="","F22"="","M02"="","M04"="",
                    "M03"="","M05"="","M08"="","M12"="","M14"="","M15"="",
                    "M17"="","_o"=" 1", "_c"=" 2", "ang"="angry", "calm"="calm",
                    "disg"="disgust", "exc"="excited", "fear"="fear",
                    "hap"="happy", "neut"="neutral", "sad"="sad",
                    "surp"="surprise"))

s2k_dend <- s2k %>% as.dendrogram
s2a_dend <- s2a %>% as.dendrogram

#Dendrograms
par(mfrow = c(1,2))
s2a_dend %>% set("branches_lwd", 2) %>%  set("branches_k_color",value=c(4,2,3), k=3) %>% plot(main="Adults")
s2k_dend %>% set("branches_lwd", 2) %>%  set("branches_k_color",value=c(2,4,3), k=3) %>% plot(main="Children")
```

# 6b. Similarity Indices {.tabset}

Here we provide additional indices of similarity between hierarchical cluster solutions and for all possible k values. While the manuscript primarily reported on the three cluster solution, we understand selecting the number of clusters is inexact. For this reason, we give additional similarity measures across all cluster sizes.

```{r data prep age bins hclust}
# Same Individual Sort
s1_3 <- avg_dist_byBin %>%
  dplyr::filter(sort=="Sort1"&age_bin=="3 to 4") %>%
  ungroup() %>%
  select(dist_obj)

s1_4 <- avg_dist_byBin %>%
  dplyr::filter(sort=="Sort1"&age_bin=="4 to 5") %>%
  ungroup() %>%
  select(dist_obj)

s1_5 <- avg_dist_byBin %>%
  dplyr::filter(sort=="Sort1"&age_bin=="5 to 6") %>%
  ungroup() %>%
  select(dist_obj)

s1_6 <- avg_dist_byBin %>%
  dplyr::filter(sort=="Sort1"&age_bin=="6 to 7") %>%
  ungroup() %>%
  select(dist_obj)

s1_a <- avg_dist_byBin %>%
  dplyr::filter(sort=="Sort1"&age_bin=="adults") %>%
  ungroup() %>%
  select(dist_obj)

#clean dist object
s1_3_dist <- s1_3$dist_obj[[1]]

s1_4_dist  <- s1_4$dist_obj[[1]]

s1_5_dist  <- s1_5$dist_obj[[1]]

s1_6_dist  <- s1_6$dist_obj[[1]] 

s1_a_dist  <- s1_a$dist_obj[[1]]

#hclust
s1_3 <- s1_3$dist_obj[[1]] %>%
  hclust(method = "ward.D2")

s1_4 <- s1_4$dist_obj[[1]] %>%
  hclust(method = "ward.D2")

s1_5 <- s1_5$dist_obj[[1]] %>%
  hclust(method = "ward.D2")

s1_6 <- s1_6$dist_obj[[1]] %>%
  hclust(method = "ward.D2")

s1_a <- s1_a$dist_obj[[1]] %>%
  hclust(method = "ward.D2")

s1_3[["labels"]] <- s1_3[["labels"]] %>%
  str_replace_all(c("M07"="","_o"=" 1", "_c"=" 2", "ang"="angry","calm"="calm","disg"="disgust",
                    "exc"="excited","fear"="fear", "hap"="happy",
                    "neut"="neutral","sad"="sad", "surp"="surprise"))
s1_4[["labels"]] <- s1_4[["labels"]] %>%
  str_replace_all(c("M07"="","_o"=" 1", "_c"=" 2", "ang"="angry","calm"="calm","disg"="disgust",
                    "exc"="excited","fear"="fear", "hap"="happy",
                    "neut"="neutral","sad"="sad", "surp"="surprise"))
s1_5[["labels"]] <- s1_5[["labels"]] %>%
  str_replace_all(c("M07"="","_o"=" 1", "_c"=" 2", "ang"="angry","calm"="calm","disg"="disgust",
                    "exc"="excited","fear"="fear", "hap"="happy",
                    "neut"="neutral","sad"="sad", "surp"="surprise"))
s1_6[["labels"]] <- s1_6[["labels"]] %>%
  str_replace_all(c("M07"="","_o"=" 1", "_c"=" 2", "ang"="angry","calm"="calm","disg"="disgust",
                    "exc"="excited","fear"="fear", "hap"="happy",
                    "neut"="neutral","sad"="sad", "surp"="surprise"))
s1_a[["labels"]] <- s1_a[["labels"]] %>%
  str_replace_all(c("M07"="","_o"=" 1", "_c"=" 2", "ang"="angry","calm"="calm","disg"="disgust",
                    "exc"="excited","fear"="fear", "hap"="happy",
                    "neut"="neutral","sad"="sad", "surp"="surprise"))

s1_3_dend <- s1_3 %>% as.dendrogram
s1_4_dend <- s1_4 %>% as.dendrogram
s1_5_dend <- s1_5 %>% as.dendrogram
s1_6_dend <- s1_6 %>% as.dendrogram
s1_a_dend <- s1_a %>% as.dendrogram

# Different Individuals Sort
s2_3 <- avg_dist_byBin %>%
  dplyr::filter(sort=="Sort2"&age_bin=="3 to 4") %>%
  ungroup() %>%
  select(dist_obj)

s2_4 <- avg_dist_byBin %>%
  dplyr::filter(sort=="Sort2"&age_bin=="4 to 5") %>%
  ungroup() %>%
  select(dist_obj)

s2_5 <- avg_dist_byBin %>%
  dplyr::filter(sort=="Sort2"&age_bin=="5 to 6") %>%
  ungroup() %>%
  select(dist_obj)

s2_6 <- avg_dist_byBin %>%
  dplyr::filter(sort=="Sort2"&age_bin=="6 to 7") %>%
  ungroup() %>%
  select(dist_obj)

s2_a <- avg_dist_byBin %>%
  dplyr::filter(sort=="Sort2"&age_bin=="adults") %>%
  ungroup() %>%
  select(dist_obj)

#dist objects
s2_3_dist <- s2_3$dist_obj[[1]]

s2_4_dist <- s2_4$dist_obj[[1]]

s2_5_dist <- s2_5$dist_obj[[1]]

s2_6_dist <- s2_6$dist_obj[[1]]

s2_a_dist <- s2_a$dist_obj[[1]]

#hclust objects
s2_3 <- s2_3$dist_obj[[1]] %>%
  hclust(method = "ward.D2")

s2_4 <- s2_4$dist_obj[[1]] %>%
  hclust(method = "ward.D2")

s2_5 <- s2_5$dist_obj[[1]] %>%
  hclust(method = "ward.D2")

s2_6 <- s2_6$dist_obj[[1]] %>%
  hclust(method = "ward.D2")

s2_a <- s2_a$dist_obj[[1]] %>%
  hclust(method = "ward.D2")

#clean up labels
s2_3[["labels"]] <- s2_3[["labels"]] %>%
  str_replace_all(c("F01"="","F04"="","F07"="","F10"="", "F13"="",
                    "F14"="","F15"="", "F17"="","F22"="","M02"="","M04"="",
                    "M03"="","M05"="","M08"="","M12"="","M14"="","M15"="",
                    "M17"="","_o"=" 1", "_c"=" 2", "ang"="angry", "calm"="calm",
                    "disg"="disgust", "exc"="excited", "fear"="fear",
                    "hap"="happy", "neut"="neutral", "sad"="sad",
                    "surp"="surprise"))
s2_4[["labels"]] <- s2_4[["labels"]] %>%
  str_replace_all(c("F01"="","F04"="","F07"="","F10"="", "F13"="",
                    "F14"="","F15"="", "F17"="","F22"="","M02"="","M04"="",
                    "M03"="","M05"="","M08"="","M12"="","M14"="","M15"="",
                    "M17"="","_o"=" 1", "_c"=" 2", "ang"="angry", "calm"="calm",
                    "disg"="disgust", "exc"="excited", "fear"="fear",
                    "hap"="happy", "neut"="neutral", "sad"="sad",
                    "surp"="surprise"))
s2_5[["labels"]] <- s2_5[["labels"]] %>%
  str_replace_all(c("F01"="","F04"="","F07"="","F10"="", "F13"="",
                    "F14"="","F15"="", "F17"="","F22"="","M02"="","M04"="",
                    "M03"="","M05"="","M08"="","M12"="","M14"="","M15"="",
                    "M17"="","_o"=" 1", "_c"=" 2", "ang"="angry", "calm"="calm",
                    "disg"="disgust", "exc"="excited", "fear"="fear",
                    "hap"="happy", "neut"="neutral", "sad"="sad",
                    "surp"="surprise"))
s2_6[["labels"]] <- s2_6[["labels"]] %>%
  str_replace_all(c("F01"="","F04"="","F07"="","F10"="", "F13"="",
                    "F14"="","F15"="", "F17"="","F22"="","M02"="","M04"="",
                    "M03"="","M05"="","M08"="","M12"="","M14"="","M15"="",
                    "M17"="","_o"=" 1", "_c"=" 2", "ang"="angry", "calm"="calm",
                    "disg"="disgust", "exc"="excited", "fear"="fear",
                    "hap"="happy", "neut"="neutral", "sad"="sad",
                    "surp"="surprise"))
s2_a[["labels"]] <- s2_a[["labels"]] %>%
  str_replace_all(c("F01"="","F04"="","F07"="","F10"="", "F13"="",
                    "F14"="","F15"="", "F17"="","F22"="","M02"="","M04"="",
                    "M03"="","M05"="","M08"="","M12"="","M14"="","M15"="",
                    "M17"="","_o"=" 1", "_c"=" 2", "ang"="angry", "calm"="calm",
                    "disg"="disgust", "exc"="excited", "fear"="fear",
                    "hap"="happy", "neut"="neutral", "sad"="sad",
                    "surp"="surprise"))


s2_3_dend <- s2_3 %>% as.dendrogram
s2_4_dend <- s2_4 %>% as.dendrogram
s2_5_dend <- s2_5 %>% as.dendrogram
s2_6_dend <- s2_6 %>% as.dendrogram
s2_a_dend <- s2_a %>% as.dendrogram
```

## Adjusted Rand
An adjusted Rand index of 0 indicates two clusters have a Rand index that matches the expected value for random groupings (i.e., no similiarity), with higher and lower values indicating higher- or lower-than-chance level similarity between the two clusters.

In all graphs, we are making comparisons between Adults and different groups of children for all possible values of k.

```{r, include=FALSE}
library(fossil)
```

```{r, warning=FALSE}
# Same Individual: Children versus Adults
m = c(1,2,3,4,5,6,7,8,9,10,11,12,13,14,15,16,17)
calc_AR <- function(x) {
  adj.rand.index(cutree(s1a, k=x),cutree(s1k, k=x))
}
# get adjusted rand for each cluster
ar <- map_dbl(m, calc_AR)
ar <- as.data.frame(ar)
ar <- as_tibble(rownames_to_column(ar))
ar <- ar %>% rename(cluster = rowname, adjusted_rand = ar) %>% mutate(cluster = as.numeric(cluster))
ar_plot1 <- ar %>% ggplot(mapping = aes(x = cluster, y = adjusted_rand))

ar_plot1 <- ar_plot1 + geom_point()+ geom_line(color="red")+
  xlab("k")+
  ylab("Adjusted Rand Index") +
  scale_x_continuous(breaks=c(2,3,4,5,6,7,8,9,10,11,12,13,14,15,16,17), limits=c(2,17)) +
  scale_y_continuous(limits=c(0,1), breaks=c(0,0.1,0.2,0.3,0.4,0.5,0.6,0.7,0.8,0.9,1.0)) +
  theme_linedraw() +
  ggtitle("Same Individual: Children")+
  theme(text=element_text(size=16, family="Times"), plot.title = element_text(size=16, face="bold",family="Times", hjust=.5),
        axis.title.x = element_text(size=20, face="plain",family="Times", vjust=.5),
        axis.title.y = element_text(size=20, face="plain",family="Times", vjust=.5))

#3-year-olds
m = c(1,2,3,4,5,6,7,8,9,10,11,12,13,14,15,16,17)
calc_AR <- function(x) {
  adj.rand.index(cutree(s1a, k=x),cutree(s1_3, k=x))
}
# get adjusted rand for each cluster
ar <- map_dbl(m, calc_AR)
ar <- as.data.frame(ar)
ar <- as_tibble(rownames_to_column(ar))
ar <- ar %>% rename(cluster = rowname, adjusted_rand = ar) %>% mutate(cluster = as.numeric(cluster))
ar_plot1a <- ar %>% ggplot(mapping = aes(x = cluster, y = adjusted_rand))

ar_plot1a <- ar_plot1a + geom_point()+ geom_line(color="red")+
  xlab("k")+
  ylab("Adjusted Rand Index") +
  scale_x_continuous(breaks=c(2,3,4,5,6,7,8,9,10,11,12,13,14,15,16,17), limits=c(2,17)) +
  scale_y_continuous(limits=c(0,1), breaks=c(0,0.1,0.2,0.3,0.4,0.5,0.6,0.7,0.8,0.9,1.0)) +
  theme_linedraw() +
  ggtitle("3-year-olds")+
  theme(text=element_text(size=16, family="Times"), plot.title = element_text(size=16, face="bold",family="Times", hjust=.5),
        axis.title.x = element_text(size=14, face="plain",family="Times", vjust=.5),
        axis.title.y = element_text(size=14, face="plain",family="Times", vjust=.5))

#4-year-olds
m = c(1,2,3,4,5,6,7,8,9,10,11,12,13,14,15,16,17)
calc_AR <- function(x) {
  adj.rand.index(cutree(s1a, k=x),cutree(s1_4, k=x))
}
# get adjusted rand for each cluster
ar <- map_dbl(m, calc_AR)
ar <- as.data.frame(ar)
ar <- as_tibble(rownames_to_column(ar))
ar <- ar %>% rename(cluster = rowname, adjusted_rand = ar) %>% mutate(cluster = as.numeric(cluster))
ar_plot1b <- ar %>% ggplot(mapping = aes(x = cluster, y = adjusted_rand))

ar_plot1b <- ar_plot1b + geom_point()+ geom_line(color="red")+
  xlab("k")+
  ylab("Adjusted Rand Index") +
  scale_x_continuous(breaks=c(2,3,4,5,6,7,8,9,10,11,12,13,14,15,16,17), limits=c(2,17)) +
  scale_y_continuous(limits=c(0,1), breaks=c(0,0.1,0.2,0.3,0.4,0.5,0.6,0.7,0.8,0.9,1.0)) +
  theme_linedraw() +
  ggtitle("4-year-olds")+
  theme(text=element_text(size=16, family="Times"), plot.title = element_text(size=16, face="bold",family="Times", hjust=.5),
        axis.title.x = element_text(size=14, face="plain",family="Times", vjust=.5),
        axis.title.y = element_text(size=14, face="plain",family="Times", vjust=.5))

#5-year-olds
m = c(1,2,3,4,5,6,7,8,9,10,11,12,13,14,15,16,17)
calc_AR <- function(x) {
  adj.rand.index(cutree(s1a, k=x),cutree(s1_5, k=x))
}
# get adjusted rand for each cluster
ar <- map_dbl(m, calc_AR)
ar <- as.data.frame(ar)
ar <- as_tibble(rownames_to_column(ar))
ar <- ar %>% rename(cluster = rowname, adjusted_rand = ar) %>% mutate(cluster = as.numeric(cluster))
ar_plot1c <- ar %>% ggplot(mapping = aes(x = cluster, y = adjusted_rand))

ar_plot1c <- ar_plot1c + geom_point()+ geom_line(color="red")+
  xlab("k")+
  ylab("Adjusted Rand Index") +
  scale_x_continuous(breaks=c(2,3,4,5,6,7,8,9,10,11,12,13,14,15,16,17), limits=c(2,17)) +
  scale_y_continuous(limits=c(0,1), breaks=c(0,0.1,0.2,0.3,0.4,0.5,0.6,0.7,0.8,0.9,1.0)) +
  theme_linedraw() +
  ggtitle("5-year-olds")+
  theme(text=element_text(size=16, family="Times"), plot.title = element_text(size=16, face="bold",family="Times", hjust=.5),
        axis.title.x = element_text(size=14, face="plain",family="Times", vjust=.5),
        axis.title.y = element_text(size=14, face="plain",family="Times", vjust=.5))

#6-year-olds
m = c(1,2,3,4,5,6,7,8,9,10,11,12,13,14,15,16,17)
calc_AR <- function(x) {
  adj.rand.index(cutree(s1a, k=x),cutree(s1_6, k=x))
}
# get adjusted rand for each cluster
ar <- map_dbl(m, calc_AR)
ar <- as.data.frame(ar)
ar <- as_tibble(rownames_to_column(ar))
ar <- ar %>% rename(cluster = rowname, adjusted_rand = ar) %>% mutate(cluster = as.numeric(cluster))
ar_plot1d <- ar %>% ggplot(mapping = aes(x = cluster, y = adjusted_rand))

ar_plot1d <- ar_plot1d + geom_point()+ geom_line(color="red")+
  xlab("k")+
  ylab("Adjusted Rand Index") +
  scale_x_continuous(breaks=c(2,3,4,5,6,7,8,9,10,11,12,13,14,15,16,17), limits=c(2,17)) +
  scale_y_continuous(limits=c(0,1), breaks=c(0,0.1,0.2,0.3,0.4,0.5,0.6,0.7,0.8,0.9,1.0)) +
  theme_linedraw() +
  ggtitle("6-year-olds")+
  theme(text=element_text(size=16, family="Times"), plot.title = element_text(size=16, face="bold",family="Times", hjust=.5),
        axis.title.x = element_text(size=14, face="plain",family="Times", vjust=.5),
        axis.title.y = element_text(size=14, face="plain",family="Times", vjust=.5))

# Diff Individual: Children versus Adults
calc_AR <- function(x) {
  adj.rand.index(cutree(s2a, k=x),cutree(s2k, k=x))
}

ar <- map_dbl(m, calc_AR)
ar <- as.data.frame(ar)
ar <- as_tibble(rownames_to_column(ar))
ar <- ar %>% rename(cluster = rowname, adjusted_rand = ar) %>% mutate(cluster = as.numeric(cluster))
ar_plot2 <- ar %>% ggplot(mapping = aes(x = cluster, y = adjusted_rand))

ar_plot2 <- ar_plot2 + geom_point()+ geom_line(color="red")+
  xlab("k")+
  ylab("Adjusted Rand Index") +
  scale_x_continuous(breaks=c(2,3,4,5,6,7,8,9,10,11,12,13,14,15,16,17), limits=c(2,17)) +
  scale_y_continuous(limits=c(0,1), breaks=c(0,0.1,0.2,0.3,0.4,0.5,0.6,0.7,0.8,0.9,1.0)) +
  theme_linedraw() +
  ggtitle("Diff. Individuals: Children")+
  theme(text=element_text(size=16, family="Times"), plot.title = element_text(size=16, face="bold",family="Times", hjust=.5),
        axis.title.x = element_text(size=20, face="plain",family="Times", vjust=.5),
        axis.title.y = element_text(size=20, face="plain",family="Times", vjust=.5))

#3-year-olds
m = c(1,2,3,4,5,6,7,8,9,10,11,12,13,14,15,16,17)
calc_AR <- function(x) {
  adj.rand.index(cutree(s2a, k=x),cutree(s2_3, k=x))
}
# get adjusted rand for each cluster
ar <- map_dbl(m, calc_AR)
ar <- as.data.frame(ar)
ar <- as_tibble(rownames_to_column(ar))
ar <- ar %>% rename(cluster = rowname, adjusted_rand = ar) %>% mutate(cluster = as.numeric(cluster))
ar_plot2a <- ar %>% ggplot(mapping = aes(x = cluster, y = adjusted_rand))

ar_plot2a <- ar_plot2a + geom_point()+ geom_line(color="red")+
  xlab("k")+
  ylab("Adjusted Rand Index") +
  scale_x_continuous(breaks=c(2,3,4,5,6,7,8,9,10,11,12,13,14,15,16,17), limits=c(2,17)) +
  scale_y_continuous(limits=c(0,1), breaks=c(0,0.1,0.2,0.3,0.4,0.5,0.6,0.7,0.8,0.9,1.0)) +
  theme_linedraw() +
  ggtitle("3-year-olds")+
  theme(text=element_text(size=16, family="Times"), plot.title = element_text(size=16, face="bold",family="Times", hjust=.5),
        axis.title.x = element_text(size=14, face="plain",family="Times", vjust=.5),
        axis.title.y = element_text(size=14, face="plain",family="Times", vjust=.5))

#4-year-olds
m = c(1,2,3,4,5,6,7,8,9,10,11,12,13,14,15,16,17)
calc_AR <- function(x) {
  adj.rand.index(cutree(s2a, k=x),cutree(s2_4, k=x))
}
# get adjusted rand for each cluster
ar <- map_dbl(m, calc_AR)
ar <- as.data.frame(ar)
ar <- as_tibble(rownames_to_column(ar))
ar <- ar %>% rename(cluster = rowname, adjusted_rand = ar) %>% mutate(cluster = as.numeric(cluster))
ar_plot2b <- ar %>% ggplot(mapping = aes(x = cluster, y = adjusted_rand))

ar_plot2b <- ar_plot2b + geom_point()+ geom_line(color="red")+
  xlab("k")+
  ylab("Adjusted Rand Index") +
  scale_x_continuous(breaks=c(2,3,4,5,6,7,8,9,10,11,12,13,14,15,16,17), limits=c(2,17)) +
  scale_y_continuous(limits=c(0,1), breaks=c(0,0.1,0.2,0.3,0.4,0.5,0.6,0.7,0.8,0.9,1.0)) +
  theme_linedraw() +
  ggtitle("4-year-olds")+
  theme(text=element_text(size=16, family="Times"), plot.title = element_text(size=16, face="bold",family="Times", hjust=.5),
        axis.title.x = element_text(size=14, face="plain",family="Times", vjust=.5),
        axis.title.y = element_text(size=14, face="plain",family="Times", vjust=.5))

#5-year-olds
m = c(1,2,3,4,5,6,7,8,9,10,11,12,13,14,15,16,17)
calc_AR <- function(x) {
  adj.rand.index(cutree(s2a, k=x),cutree(s2_5, k=x))
}
# get adjusted rand for each cluster
ar <- map_dbl(m, calc_AR)
ar <- as.data.frame(ar)
ar <- as_tibble(rownames_to_column(ar))
ar <- ar %>% rename(cluster = rowname, adjusted_rand = ar) %>% mutate(cluster = as.numeric(cluster))
ar_plot2c <- ar %>% ggplot(mapping = aes(x = cluster, y = adjusted_rand))

ar_plot2c <- ar_plot2c + geom_point()+ geom_line(color="red")+
  xlab("k")+
  ylab("Adjusted Rand Index") +
  scale_x_continuous(breaks=c(2,3,4,5,6,7,8,9,10,11,12,13,14,15,16,17), limits=c(2,17)) +
  scale_y_continuous(limits=c(0,1), breaks=c(0,0.1,0.2,0.3,0.4,0.5,0.6,0.7,0.8,0.9,1.0)) +
  theme_linedraw() +
  ggtitle("5-year-olds")+
  theme(text=element_text(size=16, family="Times"), plot.title = element_text(size=16, face="bold",family="Times", hjust=.5),
        axis.title.x = element_text(size=14, face="plain",family="Times", vjust=.5),
        axis.title.y = element_text(size=14, face="plain",family="Times", vjust=.5))

#6-year-olds
m = c(1,2,3,4,5,6,7,8,9,10,11,12,13,14,15,16,17)
calc_AR <- function(x) {
  adj.rand.index(cutree(s2a, k=x),cutree(s2_6, k=x))
}
# get adjusted rand for each cluster
ar <- map_dbl(m, calc_AR)
ar <- as.data.frame(ar)
ar <- as_tibble(rownames_to_column(ar))
ar <- ar %>% rename(cluster = rowname, adjusted_rand = ar) %>% mutate(cluster = as.numeric(cluster))
ar_plot2d <- ar %>% ggplot(mapping = aes(x = cluster, y = adjusted_rand))

ar_plot2d <- ar_plot2d + geom_point()+ geom_line(color="red")+
  xlab("k")+
  ylab("Adjusted Rand Index") +
  scale_x_continuous(breaks=c(2,3,4,5,6,7,8,9,10,11,12,13,14,15,16,17), limits=c(2,17)) +
  scale_y_continuous(limits=c(0,1), breaks=c(0,0.1,0.2,0.3,0.4,0.5,0.6,0.7,0.8,0.9,1.0)) +
  theme_linedraw() +
  ggtitle("6-year-olds")+
  theme(text=element_text(size=16, family="Times"), plot.title = element_text(size=16, face="bold",family="Times", hjust=.5),
        axis.title.x = element_text(size=14, face="plain",family="Times", vjust=.5),
        axis.title.y = element_text(size=14, face="plain",family="Times", vjust=.5))


# All Children
plot_grid(ar_plot1,ar_plot2)

# Same Individual Sort
title1 <- ggdraw() +
  draw_label("Same Individual Sort",
    fontface = 'bold',
    x = 0,
    hjust = 0) +
  theme(plot.margin = margin(0, 0, 0, 7))
plot_row1 <- plot_grid(ar_plot1a,ar_plot1b,ar_plot1c,ar_plot1d)
plot_grid(
  title1, plot_row1,
  ncol = 1,
  rel_heights = c(0.1, 1)
)

# Different Individuals Sort
title2 <- ggdraw() +
  draw_label("Diff. Individuals Sort",
    fontface = 'bold',
    x = 0,
    hjust = 0) +
  theme(plot.margin = margin(0, 0, 0, 7))
plot_row2 <- plot_grid(ar_plot2a,ar_plot2b,ar_plot2c,ar_plot2d)
plot_grid(
  title2, plot_row2,
  ncol = 1,
  rel_heights = c(0.1, 1)
)
```

## FM Index
The Fowlkes-Mallows (FM) Index measures the similarity between two hierarchical clusterings (Fowlkes & Mallow, 1983). It ranges from 0 (no similarity) to 1 (perfect similarity). The index aims to capture how likely it is that two clusters contain the same faces. Below we plot the FM-Index for all possible cluster solutions. Within this plot, the dashed black line indicates the null hypothesis (the similarity expected if the two dendrograms are “not similar”), the red line is the rejection line, above which we can say that the two dendrograms are similar, and the dotted black line is the FM-Index that we observe when comparing the children’s hierarchical clustering solution to that of adults. In all graphs, we are making comparisons between Adults and different groups of children.


```{r fig.width=10}
# Same Individual Sort:
par(mfrow = c(2,3))
Bk_plot(s1k, s1a, main = "Same Individual Sort \nAll Children")
Bk_plot(s1a, s1_3, main = "Same Individual Sort \n3-year-olds")
Bk_plot(s1a, s1_4, main = "Same Individual Sort \n4-year-olds")
Bk_plot(s1a, s1_5, main = "Same Individual Sort \n5-year-olds")
Bk_plot(s1a, s1_6, main = "Same Individual Sort \n6-year-olds")

# Different Individuals Sort
par(mfrow = c(2,3))
Bk_plot(s2a, s2k, main = "Diff. Individual Sort \nAll Children")
Bk_plot(s2a, s2_3, main = "Diff. Individual Sort \n3-year-olds")
Bk_plot(s2a, s2_4, main = "Diff. Individual Sort \n4-year-olds")
Bk_plot(s2a, s2_5, main = "Diff. Individual Sort \n5-year-olds")
Bk_plot(s2a, s2_6, main = "Diff. Individual Sort \n6-year-olds")
```

## Entanglement
Entanglement measures whether the labels at the bottom of two dendrograms are aligned with one another. A value of 0 indicates perfect alignment, while a value of 1 indicates no alignment. To optimize alignment, we first used the untangle method. For all values, we are making comparisons between Adults and different groups of children.
```{r warning=FALSE,message=FALSE}
library(gt)
tang_tracker1 <- tibble(Age_Group = character(), entanglement= numeric())
tang_tracker2 <- tibble(Age_Group = character(), entanglement= numeric())

# Same Individual Sort
set.seed(51314)

p0 <- dendlist("Adults" = s1_a_dend, "Children" = s1k_dend)
p1 <- dendlist("Adults" = s1_a_dend, "Kids" = s1_3_dend)
p2 <- dendlist("Adults" = s1_a_dend, "Kids" = s1_4_dend)
p3 <- dendlist("Adults" = s1_a_dend, "Kids" = s1_5_dend)
p4 <- dendlist("Adults" = s1_a_dend, "Kids" = s1_6_dend)

x_untangle0 <- p0 %>% untangle(method = "random", R=1000)
x_untangle1 <- p1 %>% untangle(method = "random", R=1000)
x_untangle2 <- p2 %>% untangle(method = "random", R=1000)
x_untangle3 <- p3 %>% untangle(method = "random", R=1000)
x_untangle4 <- p4 %>% untangle(method = "random", R=1000)

tang_tracker1 <- bind_rows(tang_tracker1, tibble(Age_Group = "All Children", entanglement=entanglement(x_untangle0)))
tang_tracker1 <- bind_rows(tang_tracker1, tibble(Age_Group = "3-year-olds", entanglement=entanglement(x_untangle1)))
tang_tracker1 <- bind_rows(tang_tracker1, tibble(Age_Group = "4-year-olds", entanglement=entanglement(x_untangle2)))
tang_tracker1 <- bind_rows(tang_tracker1, tibble(Age_Group = "5-year-olds", entanglement=entanglement(x_untangle3)))
tang_tracker1 <- bind_rows(tang_tracker1, tibble(Age_Group = "6-year-olds", entanglement=entanglement(x_untangle4)))


# Different Individuals Sort
set.seed(51314)

p0 <- dendlist("Adults" = s2_a_dend, "Children" = s2k_dend)
p1 <- dendlist("Adults" = s2_a_dend, "Kids" = s2_3_dend)
p2 <- dendlist("Adults" = s2_a_dend, "Kids" = s2_4_dend)
p3 <- dendlist("Adults" = s2_a_dend, "Kids" = s2_5_dend)
p4 <- dendlist("Adults" = s2_a_dend, "Kids" = s2_6_dend)

x_untangle0 <- p0 %>% untangle(method = "random", R=1000)
x_untangle1 <- p1 %>% untangle(method = "random", R=1000)
x_untangle2 <- p2 %>% untangle(method = "random", R=1000)
x_untangle3 <- p3 %>% untangle(method = "random", R=1000)
x_untangle4 <- p4 %>% untangle(method = "random", R=1000)

tang_tracker2 <- bind_rows(tang_tracker2, tibble(Age_Group = "All Children", entanglement=entanglement(x_untangle0)))
tang_tracker2 <- bind_rows(tang_tracker2, tibble(Age_Group = "3-year-olds", entanglement=entanglement(x_untangle1)))
tang_tracker2 <- bind_rows(tang_tracker2, tibble(Age_Group = "4-year-olds", entanglement=entanglement(x_untangle2)))
tang_tracker2 <- bind_rows(tang_tracker2, tibble(Age_Group = "5-year-olds", entanglement=entanglement(x_untangle3)))
tang_tracker2 <- bind_rows(tang_tracker2, tibble(Age_Group = "6-year-olds", entanglement=entanglement(x_untangle4)))

tab1 <- bind_cols(tang_tracker1, tang_tracker2)
tab1 %>% gt(tab1) %>% cols_align(align="center") %>% cols_label(Age_Group...1 = "Age Groups", entanglement...2=md("Entanglement"),Age_Group...3 = "Age Groups", entanglement...4=md("Entanglement")) %>%   tab_header(
    title = md("Entanglement"),
    subtitle = md("**Same Individual versus Different Individuals Sort**")
  )
```

## Cophenetic Correlation
c is the cophenetic correlation coefficient between the two dendrograms. This value ranges from -1 to 1 with values near 0 suggesting that the two dendrograms are not statistically similar. For all values, we are making comparisons between Adults and different groups of children.
```{r, warning=FALSE, message=FALSE}

cor_tracker1 <- tibble(Age_Group = character(),c = numeric())
cor_tracker2 <- tibble(Age_Group = character(),c = numeric())

#Same Individual Sort (Comparison to Adults)
cor_tracker1 <- bind_rows(cor_tracker1, tibble(Age_Group = "All Children", c=cor_cophenetic(s1a, s1k)))
cor_tracker1 <- bind_rows(cor_tracker1, tibble(Age_Group = "3-year-olds", c=cor_cophenetic(s1a, s1_3)))
cor_tracker1 <- bind_rows(cor_tracker1, tibble(Age_Group = "4-year-olds", c=cor_cophenetic(s1a, s1_4)))
cor_tracker1 <- bind_rows(cor_tracker1, tibble(Age_Group = "5-year-olds", c=cor_cophenetic(s1a, s1_5)))
cor_tracker1 <- bind_rows(cor_tracker1, tibble(Age_Group = "6-year-olds", c=cor_cophenetic(s1a, s1_6)))

#Different Individuals Sort (Comparison to Adults)
cor_tracker2 <- bind_rows(cor_tracker2, tibble(Age_Group = "All Children", c=cor_cophenetic(s2a, s2k)))
cor_tracker2 <- bind_rows(cor_tracker2, tibble(Age_Group = "3-year-olds", c=cor_cophenetic(s2a, s2_3)))
cor_tracker2 <- bind_rows(cor_tracker2, tibble(Age_Group = "4-year-olds", c=cor_cophenetic(s2a, s2_4)))
cor_tracker2 <- bind_rows(cor_tracker2, tibble(Age_Group = "5-year-olds", c=cor_cophenetic(s2a, s2_5)))
cor_tracker2 <- bind_rows(cor_tracker2, tibble(Age_Group = "6-year-olds", c=cor_cophenetic(s2a, s2_6)))

tab1 <- bind_cols(cor_tracker1, cor_tracker2)
tab1 %>% gt(tab1) %>% cols_align(align="center") %>% cols_label(Age_Group...1 = "Age Groups", c...2=md("*c*"),Age_Group...3 = "Age Groups", c...4=md("*c*")) %>%   tab_header(
    title = md("Cophenetic Correlation"),
    subtitle = md("**Same Individual versus Different Individuals Sort**")
  )
```

## Baker's Gamma
Baker’s Gamma Index is a measure of similarity between two dendrograms. This value ranges from -1 to 1 with values near 0 suggesting that the two dendrograms are not statistically similar. For all values, we are making comparisons between Adults and different groups of children.
```{r, warning=FALSE, message=FALSE}

cor_tracker1 <- tibble(Age_Group = character(), gamma = numeric())
cor_tracker2 <- tibble(Age_Group = character(), gamma = numeric())

#Same Individual Sort (Comparison to Adults)
cor_tracker1 <- bind_rows(cor_tracker1, tibble(Age_Group = "All Children", gamma=cor_bakers_gamma(s1a, s1k)))
cor_tracker1 <- bind_rows(cor_tracker1, tibble(Age_Group = "3-year-olds", gamma=cor_bakers_gamma(s1a, s1_3)))
cor_tracker1 <- bind_rows(cor_tracker1, tibble(Age_Group = "4-year-olds", gamma=cor_bakers_gamma(s1a, s1_4)))
cor_tracker1 <- bind_rows(cor_tracker1, tibble(Age_Group = "5-year-olds", gamma=cor_bakers_gamma(s1a, s1_5)))
cor_tracker1 <- bind_rows(cor_tracker1, tibble(Age_Group = "6-year-olds", gamma=cor_bakers_gamma(s1a, s1_6)))

#Different Individuals Sort (Comparison to Adults)
cor_tracker2 <- bind_rows(cor_tracker2, tibble(Age_Group = "All Children", gamma=cor_bakers_gamma(s2a, s2k)))
cor_tracker2 <- bind_rows(cor_tracker2, tibble(Age_Group = "3-year-olds", gamma=cor_bakers_gamma(s2a, s2_3)))
cor_tracker2 <- bind_rows(cor_tracker2, tibble(Age_Group = "4-year-olds", gamma=cor_bakers_gamma(s2a, s2_4)))
cor_tracker2 <- bind_rows(cor_tracker2, tibble(Age_Group = "5-year-olds", gamma=cor_bakers_gamma(s2a, s2_5)))
cor_tracker2 <- bind_rows(cor_tracker2, tibble(Age_Group = "6-year-olds", gamma=cor_bakers_gamma(s2a, s2_6)))

tab1 <- bind_cols(cor_tracker1, cor_tracker2)
tab1 %>% gt(tab1) %>% cols_align(align="center") %>% cols_label(Age_Group...1 = "Age Groups", gamma...2=md("Baker's Gamma"),Age_Group...3 = "Age Groups", gamma...4=md("Baker's Gamma")) %>%   tab_header(
    title = md("Baker's Gamma"),
    subtitle = md("**Same Individual versus Different Individuals Sort**")
  )
```

# 6c. Hierarchical Clustering: Selecting method and k {.tabset}


## Method: Same Individual Sort
To select our hierarchical clustering method (average, single, complete, or Ward’s), we analyzed the agglomerative coefficient (AC). AC gives a measure of the strength of the clustering structure, and we found that Ward’s method gave the best solution. Furthermore, Ward’s is a generally preferred method for agglomerative (bottom-up) hierarchical clustering (Kaufman & Rousseeuw, 2009; Boehmke & Greenwell, 2020).
```{r warning=FALSE,message=FALSE}

#methods
m <- c( "average", "single", "complete", "ward")

# Different Age Groups
names(m) <- c( "average", "single", "complete", "ward")
ac <- function(x) {
  agnes(s1a_dist, method = x)$ac
}
t1 <- as.data.frame(map_dbl(m, ac)) %>% rownames_to_column() %>% rename(method="rowname",AC=`map_dbl(m, ac)`)

names(m) <- c( "average", "single", "complete", "ward")
ac <- function(x) {
  agnes(s1k_dist, method = x)$ac
}
t2 <- as.data.frame(map_dbl(m, ac)) %>% rownames_to_column() %>% rename(method="rowname",AC=`map_dbl(m, ac)`)

names(m) <- c( "average", "single", "complete", "ward")
ac <- function(x) {
  agnes(s1_3_dist, method = x)$ac
}
t3 <- as.data.frame(map_dbl(m, ac)) %>% rownames_to_column() %>% rename(method="rowname",AC=`map_dbl(m, ac)`)

names(m) <- c( "average", "single", "complete", "ward")
ac <- function(x) {
  agnes(s1_4_dist, method = x)$ac
}
t4 <- as.data.frame(map_dbl(m, ac)) %>% rownames_to_column() %>% rename(method="rowname",AC=`map_dbl(m, ac)`)

names(m) <- c( "average", "single", "complete", "ward")
ac <- function(x) {
  agnes(s1_5_dist, method = x)$ac
}
t5 <- as.data.frame(map_dbl(m, ac)) %>% rownames_to_column() %>% rename(method="rowname",AC=`map_dbl(m, ac)`)

names(m) <- c( "average", "single", "complete", "ward")
ac <- function(x) {
  agnes(s1_6_dist, method = x)$ac
}
t6 <- as.data.frame(map_dbl(m, ac)) %>% rownames_to_column() %>% rename(method="rowname",AC=`map_dbl(m, ac)`)

tab1 <- bind_cols(t1,t2,t3,t4,t5,t6)
gt(tab1) %>% cols_align(align="center") %>% tab_header(title = md("Selecting Hierarchical Clustering Method"), subtitle = md("**Same Individual Sort**"))%>% cols_align(align="center") %>% cols_label(method...1 = "Adults", AC...2=md("*AC*"), method...3 = "Children", AC...4=md("*AC*"),method...5 = "3-years",
AC...6=md("*AC*"),method...7 = "4-years", 
AC...8=md("*AC*"),method...9 = "5-years", AC...10=md("*AC*"),method...11 = "6-years", 
AC...12=md("*AC*")) 
```


## Method: Different Individual Sort
To select our hierarchical clustering method (average, single, complete, or Ward’s), we analyzed the agglomerative coefficient (AC). AC gives a measure of the strength of the clustering structure, and we found that Ward’s method gave the best solution. Furthermore, Ward’s is a generally preferred method for agglomerative (bottom-up) hierarchical clustering (Kaufman & Rousseeuw, 2009; Boehmke & Greenwell, 2020).
```{r warning=FALSE,message=FALSE}
#methods
m <- c( "average", "single", "complete", "ward")

# Different Age Groups
names(m) <- c( "average", "single", "complete", "ward")
ac <- function(x) {
  agnes(s2a_dist, method = x)$ac
}
t1 <- as.data.frame(map_dbl(m, ac)) %>% rownames_to_column() %>% rename(method="rowname",AC=`map_dbl(m, ac)`)

names(m) <- c( "average", "single", "complete", "ward")
ac <- function(x) {
  agnes(s2k_dist, method = x)$ac
}
t2 <- as.data.frame(map_dbl(m, ac)) %>% rownames_to_column() %>% rename(method="rowname",AC=`map_dbl(m, ac)`)

names(m) <- c( "average", "single", "complete", "ward")
ac <- function(x) {
  agnes(s2_3_dist, method = x)$ac
}
t3 <- as.data.frame(map_dbl(m, ac)) %>% rownames_to_column() %>% rename(method="rowname",AC=`map_dbl(m, ac)`)

names(m) <- c( "average", "single", "complete", "ward")
ac <- function(x) {
  agnes(s2_4_dist, method = x)$ac
}
t4 <- as.data.frame(map_dbl(m, ac)) %>% rownames_to_column() %>% rename(method="rowname",AC=`map_dbl(m, ac)`)

names(m) <- c( "average", "single", "complete", "ward")
ac <- function(x) {
  agnes(s2_5_dist, method = x)$ac
}
t5 <- as.data.frame(map_dbl(m, ac)) %>% rownames_to_column() %>% rename(method="rowname",AC=`map_dbl(m, ac)`)

names(m) <- c( "average", "single", "complete", "ward")
ac <- function(x) {
  agnes(s2_6_dist, method = x)$ac
}
t6 <- as.data.frame(map_dbl(m, ac)) %>% rownames_to_column() %>% rename(method="rowname",AC=`map_dbl(m, ac)`)

tab1 <- bind_cols(t1,t2,t3,t4,t5,t6)
gt(tab1) %>% cols_align(align="center") %>% tab_header(title = md("Selecting Hierarchical Clustering Method"), subtitle = md("**Different Individuals Sort**"))%>% cols_align(align="center") %>% cols_label(method...1 = "Adults", AC...2=md("*AC*"), method...3 = "Children", AC...4=md("*AC*"),method...5 = "3-years",
AC...6=md("*AC*"),method...7 = "4-years", 
AC...8=md("*AC*"),method...9 = "5-years", AC...10=md("*AC*"),method...11 = "6-years", 
AC...12=md("*AC*")) 
```

## Selecting k: Same Individual Sort
As we used hierarchical clustering and presented dendrograms, we displayed all possible clusters in the manuscript. However, we did chose to highlight the three cluster solution (k=3) in the dendrograms we displayed. Below we present bar plots of the height of the different clusters in the dendrogram. Large changes in these heights can be helpful for determining the optimal number of clusters to display. Through observations of these graphs, subjective evaluation of the dendrograms, patterns of valence among the images, and a desire to be consistent across different age ranges, we opted to display the three cluster solution. However, for full transparency we also report similarity indices for all values of k in section 6b (above).
```{r, warning =FALSE}

s1_cluster <- as.data.frame(s1a$height) %>% rename(adults="s1a$height") %>% mutate(children=s1k$height,three_years=s1_3$height,four_years=s1_4$height,five_years=s1_5$height,six_years=s1_6$height) %>% rownames_to_column() %>% rename(cluster=rowname) %>% mutate(cluster=abs(as.numeric(cluster)-18))

s1_cluster_dL <- s1_cluster %>% group_by(cluster) %>% gather(key = age_group, value = height, adults,children,three_years,four_years,five_years,six_years)

s1_cluster_dL <- s1_cluster_dL %>% mutate(age_group=case_when(
  age_group == "adults" ~ "adults",
  age_group =="children" ~ "children",
  age_group =="three_years" ~ "3-year-olds",
  age_group =="four_years" ~ "4-year-olds",
  age_group =="five_years" ~ "5-year-olds",
  age_group =="six_years" ~ "6-year-olds"))

ggplot(data=s1_cluster_dL,aes(x=cluster,y=height, fill=age_group)) + geom_bar(stat='identity', color="black") + facet_wrap(~age_group) +
  theme(legend.position="none")+
  scale_x_continuous(breaks=c(1,3,5,7,9,11,13,15,17))+
  ggtitle("Height of dendrogram for cluster k")
  


```



## Selecting k: Different Individuals Sort
As we used hierarchical clustering and presented dendrograms, we displayed all possible clusters in the manuscript. However, we did chose to highlight the three cluster solution (k=3) in the dendrograms we displayed. Below we present bar plots of the height of the different clusters in the dendrogram. Large changes in these heights can be helpful for determining the optimal number of clusters to display. Through observations of these graphs, subjective evaluation of the dendrograms, patterns of valence among the images, and a desire to be consistent across different age ranges, we opted to display the three cluster solution. However, for full transparency we also report similarity indices for all values of k in section 6b (above).
```{r, warning=FALSE}
s2_cluster <- as.data.frame(s2a$height) %>% rename(adults="s2a$height") %>% mutate(children=s2k$height,three_years=s2_3$height,four_years=s2_4$height,five_years=s2_5$height,six_years=s2_6$height) %>% rownames_to_column() %>% rename(cluster=rowname) %>% mutate(cluster=abs(as.numeric(cluster)-18))

s2_cluster_dL <- s2_cluster %>% group_by(cluster) %>% gather(key = age_group, value = height, adults,children,three_years,four_years,five_years,six_years)

s2_cluster_dL <- s2_cluster_dL %>% mutate(age_group=case_when(
  age_group == "adults" ~ "adults",
  age_group =="children" ~ "children",
  age_group =="three_years" ~ "3-year-olds",
  age_group =="four_years" ~ "4-year-olds",
  age_group =="five_years" ~ "5-year-olds",
  age_group =="six_years" ~ "6-year-olds"))

ggplot(data=s2_cluster_dL,aes(x=cluster,y=height, fill=age_group)) + geom_bar(stat='identity', color="black") + facet_wrap(~age_group) +
  theme(legend.position="none") +
  scale_x_continuous(breaks=c(1,3,5,7,9,11,13,15,17))+
  ggtitle("Height of dendrogram for cluster k")
  

```



# 7. Rating Data {.tabset}
Here we present more information about the ratings for all of the different images. Valence and arousal were rated on 7-point Likert scales. Positivity and Negativity where rated on a 4-point scale using the evaluative space grid. 
```{r rating data prep}
library(gmodels)
raw_ratings <- read.csv(here(root_path,"analysis","paper_2020","processed_data","ratings_raw.csv"))

ratings_dL <- raw_ratings %>% group_by(participant,image) %>% gather(key = rating_type, value = response, arousal,valence,negativity,positivity,negativity)
```

## Ratings: Same Individual Sort
```{r, warning=FALSE}
r1 <- ratings_dL %>%
  filter(grepl('M07',image)) %>%
  group_by(image, rating_type) %>%
  summarise(rating=ci(response)[1], 
            lowCI = ci(response)[2],
            hiCI = ci(response)[3], 
            sd = ci (response)[4]) %>%
  ggplot(aes(x=image,y=rating)) + 
  geom_bar(stat="identity") +
  geom_errorbar(aes(ymin=lowCI,ymax=hiCI),width=0,color="black")+
  scale_fill_brewer(palette="Set1")+
  facet_wrap(~rating_type)+
  theme(legend.position="none")+
  ggtitle("Same Individual Ratings") +
  theme(axis.text.x = element_text(angle = 90, hjust = 1,vjust=0.5))

r1
```


## Ratings: Different Individuals Sort
```{r, warning=FALSE}
r2 <- ratings_dL %>%
  filter(!grepl('M07',image)) %>%
  group_by(image, rating_type) %>%
  summarise(rating=ci(response)[1], 
            lowCI = ci(response)[2],
            hiCI = ci(response)[3], 
            sd = ci (response)[4]) %>%
  ggplot(aes(x=image,y=rating)) + 
  geom_bar(stat="identity") +
  geom_errorbar(aes(ymin=lowCI,ymax=hiCI),width=0,color="black")+
  scale_fill_brewer(palette="Set1")+
  facet_wrap(~rating_type)+
  theme(legend.position="none")+
  ggtitle("Different Individual Ratings") +
  theme(axis.text.x = element_text(angle = 90, hjust = 1,vjust=0.5))

r2
```

## Range of responses
```{r}
ratings_dL %>%
  ggplot(aes(x=response,fill=rating_type)) + 
  geom_bar(width = 0.5, color="black") +
  facet_wrap(~rating_type, strip.position="bottom") +
  theme(legend.position="none") +
  ggtitle("Range of rating responses")
```

## Correlation Tables of Dimension Ratings
```{r}

cor1 <- ratings_pairs %>% filter(sort=="Sort1") %>% select(dist_valence,dist_arousal,dist_pos,dist_neg) %>% cor()
cor2 <- ratings_pairs %>% filter(sort=="Sort2") %>% select(dist_valence,dist_arousal,dist_pos,dist_neg) %>% cor()

cor1 <- as.data.frame(cor1) %>% rownames_to_column() %>% mutate(rowname=case_when(
  rowname=="dist_valence" ~ "Valence",
  rowname=="dist_arousal" ~ "Arousal",
  rowname=="dist_pos" ~ "Positivity",
  rowname=="dist_neg" ~ "Negativity")) %>% rename(" "=rowname)
cor1 %>% gt(cor1) %>% cols_align(align="center") %>% cols_label(dist_valence =md("Valence"), dist_arousal=md("Arousal"),dist_pos = md("Positivity"), dist_neg=md("Negativity")) %>%   
  tab_header(title = md("Same Individual Sort"))

cor2 <- as.data.frame(cor2) %>% rownames_to_column() %>% mutate(rowname=case_when(
  rowname=="dist_valence" ~ "Valence",
  rowname=="dist_arousal" ~ "Arousal",
  rowname=="dist_pos" ~ "Positivity",
  rowname=="dist_neg" ~ "Negativity")) %>% rename(" "=rowname)
cor2 %>% gt(cor2) %>% cols_align(align="center") %>% cols_label(dist_valence =md("Valence"), dist_arousal=md("Arousal"),dist_pos = md("Positivity"), dist_neg=md("Negativity")) %>%   
  tab_header(title = md("Different Individual Sort"))

```

# Session Info

```{r}
sessionInfo()
```