---
title: "Grid: Hierarchical Clustering"
output:
  html_document:
    code_folding: hide
    toc: true
    toc_depth: 3
    toc_float: true
---

# Load packages and data
```{r load packages, include=FALSE}
library(tidyverse)
library(harrietr)
library(ggdendro)
library(ggimage)
library(dendroextras)
library(dendextend)
library(corrplot)
library(ggcorrplot)
library(here)
library(cowplot)
library(car)
library(cluster)
library(factoextra)
library(fossil)
library(knitr)
theme_set(theme_cowplot())
root_path <- here()
source(here("analysis","paper_2020","Grid_helper_functions.R"))

#read in data
subj_dist_long <- read.csv(here("analysis","paper_2020","processed_data","Grid_subject_distance_item_pairs.csv"))

ratings <- read.csv(here("analysis","paper_2020","processed_data","ratings_data.csv")) %>% 
  select(image, sort, pos, neg, valence, arousal) %>%
  rename(label = image)
#clean ratings names
ratings_clean <- ratings %>% 
  mutate(clean_label = str_replace_all(label,c("F01"="","F04"="","F07"="","F10"="", "F13"="", "F14"="","F15"="", "F17"="","F22"="","M02"="","M04"="",
                    "M03"="","M07"="", "M05"="","M08"="","M12"="","M14"="","M15"="",
                    "M17"="","_o"=" 1", "_c"=" 2", "ang"="angry", "calm"="calm",
                    "disg"="disgust", "exc"="excited", "fear"="fear",
                    "hap"="happy", "neut"="neutral", "sad"="sad",
                    "surp"="surprise"))) %>%
  rename(old_label=label)
```

# Define functions for clustering

```{r}
clean_labels_sort <- function(cluster_object) {
  #clean up labels
  cluster_object[["old_label"]] <- cluster_object[["labels"]]
  cluster_object[["labels"]] <- cluster_object[["labels"]] %>%
  str_replace_all(c("M07"="","F01"="","F04"="","F07"="","F10"="", "F13"="",
                    "F14"="","F15"="", "F17"="","F22"="","M02"="","M04"="",
                    "M03"="","M05"="","M08"="","M12"="","M14"="","M15"="",
                    "M17"="","_o"=" 1", "_c"=" 2", "ang"="angry","calm"="calm","disg"="disgust",
                    "exc"="excited","fear"="fear", "hap"="happy",
                    "neut"="neutral","sad"="sad", "surp"="surprise"))
  cluster_object
}

clean_cluster <- function(d,cur_method="ward.D2",clean_labels=TRUE) {
  cur_cluster <- d %>%
    hclust(method = cur_method)
  if (clean_labels) {
    cur_cluster <- clean_labels_sort(cur_cluster)
  }
  cur_cluster
}

cut_cluster <- function(clst,cluster_num=3,add_ratings=TRUE) {
  #create labels data frame to retain old label names (useful for joining)
  labels=data.frame(old_label=pluck(clst,"old_label"),label=pluck(clst,"labels"))
  clst %>%
    cutree(cluster_num) %>%
    as.data.frame() %>%
    tibble::rownames_to_column("label") %>%
    left_join(labels) %>%
    rename(cluster3 = ".") %>%
    dplyr::left_join(ratings_clean, by = c("old_label")) %>%
    mutate(cluster3 = as.factor(cluster3))
}
```

# Group data by age groups and age bins
```{r group data, warning=FALSE}
#average across all distances
avg_dist_long_byGroup <- subj_dist_long %>%
  group_by(sort,age_group,item1,item2) %>%
  summarize(avg_dist=mean(dist)) %>%
  ungroup() %>%
  mutate(sort=as.character(sort),age_group=as.character(age_group),item1=as.character(item1),item2=as.character(item2))

#average distance object organized by sorting group
avg_dist_byGroup <- avg_dist_long_byGroup %>%
  group_by(sort,age_group) %>%
  nest() %>%
  mutate(dist_obj = purrr::map(data, long_to_dist))

#average across all distances
avg_dist_long_byBin <- subj_dist_long %>%
  group_by(sort,age_bin,item1,item2) %>%
  summarize(avg_dist=mean(dist)) %>%
  ungroup() %>%
  mutate(sort=as.character(sort),age_bin=as.character(age_bin),item1=as.character(item1),item2=as.character(item2))

#average distance object organized by sorting group
avg_dist_byBin <- avg_dist_long_byBin %>%
  group_by(sort,age_bin) %>%
  nest() %>%
  mutate(dist_obj = purrr::map(data, long_to_dist))

#### create overall grouped cluster objects ####
clusters_by_group <- avg_dist_byGroup %>%
  mutate(cluster=lapply(dist_obj, function(d) clean_cluster(d))) %>%
  mutate(dend = lapply(cluster, function(clst) clst %>% as.dendrogram())) %>%
  #cut cluster into 3 groups
  mutate(cut_cluster=lapply(cluster, function(clst) cut_cluster(clst,cluster_num=3))) %>%
  #fit model predicting valence from clusters
  mutate(model_valence = lapply(cut_cluster, function(df) if (sort=="Practice") {NA} else {lm(valence~cluster3, data=df)}))

clusters_by_bin <- avg_dist_byBin %>%
  mutate(cluster=lapply(dist_obj, function(d) clean_cluster(d))) %>%
  mutate(dend = lapply(cluster, function(clst) clst %>% as.dendrogram())) %>%
  #cut cluster into 3 groups
  mutate(cut_cluster=lapply(cluster, function(clst) cut_cluster(clst,cluster_num=3))) %>%
  #fit model predicting valence from clusters
  mutate(model_valence = lapply(cut_cluster, function(df) if (sort=="Practice") {NA} else {lm(valence~cluster3, data=df)}))
```

# Age Group (Adults v. Kids)

###  Same Individual Sort

#### Hclust

```{r}
# hclust solution
s1k <- clusters_by_group %>% filter(age_group=="kids",sort=="Sort1") %>% pull(cluster) %>% pluck(1)
s1a <- clusters_by_group %>% filter(age_group=="adults",sort=="Sort1") %>% pull(cluster) %>% pluck(1)
```

#### Does cluster group predict valence ratings?
```{r}
#children
clusters_by_group %>% filter(sort=="Sort1"&age_group=="kids") %>% pull(model_valence) %>% pluck(1) %>% Anova(type="III")

#adults
clusters_by_group %>% filter(sort=="Sort1"&age_group=="adults") %>% pull(model_valence) %>% pluck(1) %>% Anova(type="III")
```

###  Different Individuals Sort

#### Hclust

```{r}
# hclust solution
s2k <- clusters_by_group %>% filter(age_group=="kids",sort=="Sort2") %>% pull(cluster) %>% pluck(1)
s2a <- clusters_by_group %>% filter(age_group=="adults",sort=="Sort2") %>% pull(cluster) %>% pluck(1)
```

#### Does cluster group predict valence ratings?
```{r}
#children
clusters_by_group %>% filter(sort=="Sort2"&age_group=="kids") %>% pull(model_valence) %>% pluck(1) %>% Anova(type="III")

#adults
clusters_by_group %>% filter(sort=="Sort2"&age_group=="adults") %>% pull(model_valence) %>% pluck(1) %>% Anova(type="III")
```

# Age Bins

### Practice

#### Hclust
```{r}
#extract practice dendrograms
p3_dend <- clusters_by_bin %>% filter(age_bin=="3 to 4",sort=="Practice") %>% pull(dend) %>% pluck(1)
p4_dend <- clusters_by_bin %>% filter(age_bin=="4 to 5",sort=="Practice") %>% pull(dend) %>% pluck(1)
p5_dend <- clusters_by_bin %>% filter(age_bin=="5 to 6",sort=="Practice") %>% pull(dend) %>% pluck(1)
p6_dend <- clusters_by_bin %>% filter(age_bin=="6 to 7",sort=="Practice") %>% pull(dend) %>% pluck(1)
pa_dend <- clusters_by_bin %>% filter(age_bin=="adults",sort=="Practice") %>% pull(dend) %>% pluck(1)
```

#### Similarity Measures
```{r}
prac_compare <- dendlist("3-year-olds" = p3_dend, "4-year-olds" = p4_dend, "5-year-olds" = p5_dend, "6-year-olds" = p6_dend, "Adults" = pa_dend)

#Correlation
ggcorrplot(cor.dendlist(prac_compare), type = "lower", lab=TRUE, outline.color="black", ggtheme = ggplot2::theme_void())

#FM Index and AR Index
similarity_to_adults_p <- data.frame(
  sort=rep("Practice",4),
  age_bin = c("3 to 4","4 to 5","5 to 6","6 to 7"),
  fm_to_adults=c(
    Bk(pa_dend, p3_dend, k = 3)$`3`[1],
    Bk(pa_dend, p4_dend, k = 3)$`3`[1],
    Bk(pa_dend, p5_dend, k = 3)$`3`[1],
    Bk(pa_dend, p6_dend, k = 3)$`3`[1]
  ),
  ar_to_adults = c(
    adj.rand.index(cutree(pa_dend, k=3),cutree(p3_dend, k=3)),
    adj.rand.index(cutree(pa_dend, k=3),cutree(p4_dend, k=3)),
    adj.rand.index(cutree(pa_dend, k=3),cutree(p5_dend, k=3)),
    adj.rand.index(cutree(pa_dend, k=3),cutree(p6_dend, k=3))
  )
)

similarity_to_adults_p %>%
  select(age_bin,fm_to_adults,ar_to_adults)%>%
  kable(
      caption = "Comparing children's hierarchical clustering solutions to adults' during Practice (k=3)",
      col.names=c("Age Group","Fowlkes-Mallows Index","Adjusted Rand Index"),
      digits=c(0,2,2)
      )
```

###  Same Individual Sort

#### Hclust and dendrograms
```{r}
#Make dendrograms
s1_3_dend <- clusters_by_bin %>% filter(age_bin=="3 to 4",sort=="Sort1") %>% pull(dend) %>% pluck(1)
s1_4_dend <- clusters_by_bin %>% filter(age_bin=="4 to 5",sort=="Sort1") %>% pull(dend) %>% pluck(1)
s1_5_dend <- clusters_by_bin %>% filter(age_bin=="5 to 6",sort=="Sort1") %>% pull(dend) %>% pluck(1)
s1_6_dend <- clusters_by_bin %>% filter(age_bin=="6 to 7",sort=="Sort1") %>% pull(dend) %>% pluck(1)
s1_a_dend <- clusters_by_bin %>% filter(age_bin=="adults",sort=="Sort1") %>% pull(dend) %>% pluck(1)

png(filename=here(root_path,"analysis","paper_2020","plots","sort1_dendro.png"),res=300,width=2400, height=1200)
par(mfrow = c(2,3))
s1_3_dend %>% set("branches_lwd", 2) %>%  set("branches_k_color",value=c(1,1,1), k=3) %>% plot(main = "3-year-olds")
s1_4_dend %>% set("branches_lwd", 2) %>% set("branches_k_color", value=c(2,3,4), k=3) %>% plot(main = "4-year-olds")
s1_5_dend %>% set("branches_lwd", 2) %>% set("branches_k_color",value=c(2,3,4), k=3) %>% plot(main = "5-year-olds")
s1_6_dend %>% set("branches_lwd", 2) %>% set("branches_k_color",value=c(2,4,3), k=3) %>% plot(main = "6-year-olds")
s1_a_dend %>% set("branches_lwd", 2) %>% set("branches_k_color",value=c(4,3,2), k=3) %>% plot(main = "Adults")
dev.off()

```

#### Similarity Measures
```{r}
#Correlation
sort1_compare <- dendlist("3-year-olds" = s1_3_dend, "4-year-olds" = s1_4_dend, "5-year-olds" = s1_5_dend, "6-year-olds" = s1_6_dend, "Adults" = s1_a_dend)
ggcorrplot(cor.dendlist(sort1_compare), type = "lower", lab=TRUE, outline.color="black", ggtheme = ggplot2::theme_void())
ggsave(here(root_path,"analysis","paper_2020","plots","sort1_dendro_corr.png"),width=5,height=3.6)


similarity_to_adults_s1 <- data.frame(
  sort=rep("Sort1",4),
  age_bin = c("3 to 4","4 to 5","5 to 6","6 to 7"),
  fm_to_adults=c(
    Bk(s1_a_dend, s1_3_dend, k = 3)$`3`[1],
    Bk(s1_a_dend, s1_4_dend, k = 3)$`3`[1],
    Bk(s1_a_dend, s1_5_dend, k = 3)$`3`[1],
    Bk(s1_a_dend, s1_6_dend, k = 3)$`3`[1]
  ),
  ar_to_adults = c(
    adj.rand.index(cutree(s1_a_dend, k=3),cutree(s1_3_dend, k=3)),
    adj.rand.index(cutree(s1_a_dend, k=3),cutree(s1_4_dend, k=3)),
    adj.rand.index(cutree(s1_a_dend, k=3),cutree(s1_5_dend, k=3)),
    adj.rand.index(cutree(s1_a_dend, k=3),cutree(s1_6_dend, k=3))
  )
)

similarity_to_adults_s1 %>%
  select(age_bin,fm_to_adults,ar_to_adults)%>%
  kable(
      caption = "Comparing children's hierarchical clustering solutions to adults' during the Same Individual Sort (k=3)",
      col.names=c("Age Group","Fowlkes-Mallows Index","Adjusted Rand Index"),
      digits=c(0,2,2)
      )
```

#### Does cluster group predict valence ratings?
```{r}
#children
## 3 to 4
clusters_by_bin %>% filter(sort=="Sort1"&age_bin=="3 to 4") %>% pull(model_valence) %>% pluck(1) %>% Anova(type="III")

## 4 to 5
clusters_by_bin %>% filter(sort=="Sort1"&age_bin=="4 to 5") %>% pull(model_valence) %>% pluck(1) %>% Anova(type="III")

## 5 to 6
clusters_by_bin %>% filter(sort=="Sort1"&age_bin=="5 to 6") %>% pull(model_valence) %>% pluck(1) %>% Anova(type="III")

## 6 to 7
clusters_by_bin %>% filter(sort=="Sort1"&age_bin=="6 to 7") %>% pull(model_valence) %>% pluck(1) %>% Anova(type="III")

# adults
clusters_by_bin %>% filter(sort=="Sort1"&age_bin=="adults") %>% pull(model_valence) %>% pluck(1) %>% Anova(type="III")
```

###  Different Individuals Sort

#### Hclust and dendrograms
```{r}
#Make dendrograms
s2_3_dend <- clusters_by_bin %>% filter(age_bin=="3 to 4",sort=="Sort2") %>% pull(dend) %>% pluck(1)
s2_4_dend <- clusters_by_bin %>% filter(age_bin=="4 to 5",sort=="Sort2") %>% pull(dend) %>% pluck(1)
s2_5_dend <- clusters_by_bin %>% filter(age_bin=="5 to 6",sort=="Sort2") %>% pull(dend) %>% pluck(1)
s2_6_dend <- clusters_by_bin %>% filter(age_bin=="6 to 7",sort=="Sort2") %>% pull(dend) %>% pluck(1)
s2_a_dend <- clusters_by_bin %>% filter(age_bin=="adults",sort=="Sort2") %>% pull(dend) %>% pluck(1)

png(filename=here(root_path,"analysis","paper_2020","plots","sort2_dendro.png"),res=300,width=2400, height=1200)
par(mfrow = c(2,3))
s2_3_dend %>% set("branches_lwd", 2) %>%  set("branches_k_color",value=c(1,1,1), k=3) %>% plot(main = "3-year-olds")
s2_4_dend %>% set("branches_lwd", 2) %>% set("branches_k_color", value=c(2,3,4), k=3) %>% plot(main = "4-year-olds")
s2_5_dend %>% set("branches_lwd", 2) %>% set("branches_k_color",value=c(2,3,4), k=3) %>% plot(main = "5-year-olds")
s2_6_dend %>% set("branches_lwd", 2) %>% set("branches_k_color",value=c(2,4,3), k=3) %>% plot(main = "6-year-olds")
s2_a_dend %>% set("branches_lwd", 2) %>% set("branches_k_color",value=c(4,2,3), k=3) %>% plot(main = "Adults")
dev.off()
```

#### Similarity Measures
```{r}
#Correlation
sort2_compare <- dendlist("3-year-olds" = s2_3_dend, "4-year-olds" = s2_4_dend, "5-year-olds" = s2_5_dend, "6-year-olds" = s2_6_dend, "Adults" = s2_a_dend)
ggcorrplot(cor.dendlist(sort2_compare), type = "lower", lab=TRUE, outline.color="black", ggtheme = ggplot2::theme_void())
ggsave(here(root_path,"analysis","paper_2020","plots","sort2_dendro_corr.png"),width=5,height=3.6)


# FM Index and AR Index
similarity_to_adults_s2 <- data.frame(
  sort=rep("Sort2",4),
  age_bin = c("3 to 4","4 to 5","5 to 6","6 to 7"),
  fm_to_adults=c(
    Bk(s2_a_dend, s2_3_dend, k = 3)$`3`[1],
    Bk(s2_a_dend, s2_4_dend, k = 3)$`3`[1],
    Bk(s2_a_dend, s2_5_dend, k = 3)$`3`[1],
    Bk(s2_a_dend, s2_6_dend, k = 3)$`3`[1]
  ),
  ar_to_adults = c(
    adj.rand.index(cutree(s2_a_dend, k=3),cutree(s2_3_dend, k=3)),
    adj.rand.index(cutree(s2_a_dend, k=3),cutree(s2_4_dend, k=3)),
    adj.rand.index(cutree(s2_a_dend, k=3),cutree(s2_5_dend, k=3)),
    adj.rand.index(cutree(s2_a_dend, k=3),cutree(s2_6_dend, k=3))
  )
)

similarity_to_adults_s2 %>%
  select(age_bin,fm_to_adults,ar_to_adults)%>%
  kable(
      caption = "Comparing children's hierarchical clustering solutions to adults' during the Different Individual Sort (k=3)",
      col.names=c("Age Group","Fowlkes-Mallows Index","Adjusted Rand Index"),
      digits=c(0,2,2)
      )
```

#### Does cluster group predict valence ratings?
```{r}
#children
## 3 to 4
clusters_by_bin %>% filter(sort=="Sort2"&age_bin=="3 to 4") %>% pull(model_valence) %>% pluck(1) %>% Anova(type="III")

## 4 to 5
clusters_by_bin %>% filter(sort=="Sort2"&age_bin=="4 to 5") %>% pull(model_valence) %>% pluck(1) %>% Anova(type="III")

## 5 to 6
clusters_by_bin %>% filter(sort=="Sort2"&age_bin=="5 to 6") %>% pull(model_valence) %>% pluck(1) %>% Anova(type="III")

## 6 to 7
clusters_by_bin %>% filter(sort=="Sort2"&age_bin=="6 to 7") %>% pull(model_valence) %>% pluck(1) %>% Anova(type="III")

# adults
clusters_by_bin %>% filter(sort=="Sort2"&age_bin=="adults") %>% pull(model_valence) %>% pluck(1) %>% Anova(type="III")
```
